### 2023-03-25 doc/_bayes_freq.qmd
working on the method explanation on Quarto

---
editor:
    render-on-save: true
---

### Frequency uncertainty $\pi(\underline{\underline{F}} \mid \underline{\underline{R}})$ {#sec-bayes_freq}

We begin with the probability of the frequency values given the raw barcode
reads. The first assumption we make is that the inference of the frequency
values for time $t$ is independent of any other time. Therefore, we can write
the joint probability distribution as a product of independent distributions of
the form
$$
\pi(\underline{\underline{F}} \mid \underline{\underline{R}}) =
\prod_{t=1}^T \pi(\underline{f}_t \mid \underline{r}_t).
$${#eq-freq_indep}

Writing Bayes' theorem for this series of independent terms results in
$$
\pi(\underline{f}_t \mid \underline{r}_t) \propto
\pi(\underline{r}_t \mid \underline{f}_t)
\pi(\underline{f}_t).
$${#eq-bayes_freq}
To model the prior uncertainty in the value of the frequencies, we assume a very
simple model. Notice that the frequencies at any time point must satisfy the
normalization constraint
$$
\sum_{b = 1}^B f_t^{(b)} = 1, \;\; \forall \; f_t^{(b)} \in [0, 1].
$${#eq-freq_norm}
This means that the probability distribution of the $B$ frequency values lies on
the $B$-simplex, i.e., the $B$ dimensional space constrained such that the sum
of all the dimensions must add up to one. A natural selection for a probability
that satisfies the constraint in @eq-freq_norm is the so-called Dirichlet 
distribution. The Dirichlet distribution is parametrized by a vector
$$
\underline{\alpha}_t = (\alpha^{(1)}_t, \cdots, \alpha^{(B)}_t)^\dagger, 
\;\; \alpha_t^{(k)} \geq 0 \;\; \forall \; k.
$${#eq-alpha_dir}
Given these parameters, we choose as a prior distribution for the barcode
frequency values $\pi(\underline{f}_t)$ a Dirichlet distribution of the form
$$
\pi(\underline{f}_t) = \frac{
    \Gamma\left(\sum_{b=1}^B\alpha_t^{(b)}\right)
}{
    \prod_{b=1}^B\Gamma(\alpha_t^{(b)})
} 
\prod_{b=1}^B (f_t^{(b)})^{\alpha_t^{(b)}-1}.
$${#eq-dirichlet}
@eq-dirichlet can be much more compactly written as
$$
\underline{f}_t \sim \operatorname{Dirichlet}(\underline{\alpha}_t),
$${#eq-freq_prior}
where the symbol "$\sim$" is read "distributed as". We are free to choose the
value of the parameters $\underline{\alpha}_t$ according to our prior
information. A reasonable choice is to assume that before observing any data at
all, we have no reason to believe any barcode is in higher abundance compared to
any other. This might not be entirely true at the beginning of the experiment,
where we expect the neutral barcodes to be in much higher frequency compared to
the mutant barcodes. Nevertheless, this might not be true for the following time
points, making our "uniform prior over the simplex" a reasonable choice. 
Therefore, we define
$$
\alpha_t^{(b)} = 1 \;\; \forall \; b.
$${#eq-alpha_choice}

For the likelihood term $\pi(\underline{r}_t \mid \underline{f}_t)$, we assume a
very simple multinomial model. The story behind this choice goes as follows:
Given a "true" underlying barcode frequency, when the sequencing machine reads a
particular barcode, this is equivalent to having thrown a $B$-sided die, where
the face of the die indicates the barcode that is read. This die is then rolled
$R_t = \sum_b r_t^{(b)}$ times, giving the values of the reads we observe at
time $t$. Notice that this model does not account for noise that might come from
sources such as genetic drift, DNA extraction, and PCR amplification. Although
these sources of noise are obviously important, it is not clear that we have
enough statistical power to make meaningful statements about them given our data
(more on this later on). We can then compactly write the likelihood function as
$$
\underline{r}_t \mid \underline{f}_t \sim
\operatorname{Multinomial}(\underline{f}_t).
$${#eq-freq_likelihood}

This simple Multinomial-Dirichlet model for the probability of the frequency
values is analytically tractable. Specifically, the Dirichlet distribution is
the so-called *conjugate prior* to the multinomial distribution. This means the
posterior distribution $\pi(\underline{f}_t \mid \underline{r}_t)$ has a
closed-form solution of the same form as the prior $\pi(\underline{f}_t)$, but
with updated parameters. In other words, we can write (without proof) the
posterior distribution as
$$
\underline{f}_t \mid \underline{r}_t \sim
\operatorname{Dirichlet}(\underline{\beta}_t),
$${#eq-freq_posterior}
where the new parameters are of the form
$$
\beta_t^{(b)} = \alpha_t^{(b)} + r_t^{(b)}.
$$
Simply put, when the likelihood function is multinomial, and the prior 
distribution is Dirichlet, the posterior distribution is also Dirichlet, where
the parameters of the posterior are the parameters of the prior plus the number
of observations of the corresponding output. This fact will become handy
as we build the entire posterior distribution for our inference as defined in
@eq-split_posterior.