### 2023-03-25 doc/_bayes_meanfit.qmd

---
editor:
    render-on-save: true
---

### Population mean fitness uncertainty $\pi(\underline{\bar{s}}_T \mid \underline{\underline{F}}, \underline{\underline{R}})$ {#sec-bayes_meanfit}