### 2023-03-25 doc/_bayes_freq.qmd
continue working on the explanation of the model

---
editor:
    render-on-save: true
---

### Frequency uncertainty $\pi(\underline{\underline{F}} \mid \underline{\underline{R}})$ {#sec-bayes_freq}

We begin with the probability of the frequency values given the raw barcode
reads. The first assumption we make is that the inference of the frequency
values for time $t$ is independent of any other time. Therefore, we can write
the joint probability distribution as a product of independent distributions of
the form
$$
\pi(\underline{\underline{F}} \mid \underline{\underline{R}}) =
\prod_{t=1}^T \pi(\underline{f}_t \mid \underline{r}_t).
$${#eq-freq_indep}

Writing Bayes' theorem for one of these independent terms results in
$$
\pi(\underline{f}_t \mid \underline{r}_t) \propto
\pi(\underline{r}_t \mid \underline{f}_t)
\pi(\underline{f}_t).
$${#eq-bayes_freq}
To model the prior uncertainty in the value of the frequencies, we assume a very
simple model. Notice that the frequencies at any time point must satisfy the
normalization constraint
$$
\sum_{b = 1}^B f_t^{(b)} = 1, \;\; \forall \; f_t^{(b)} \in [0, 1].
$${#eq-freq_norm}
This means that the probability distribution of the $B$ frequency values lies on
the $B$-simplex, i.e., the $B$ dimensional space constrained such that the sum
of all the dimensions must add up to one. A natural selection for a probability
that satisfies the constraint in @eq-freq_norm is the so-called Dirichlet 
distribution. The Dirichlet distribution is parametrized by a vector
$$
\underline{\alpha}_t = (\alpha^{(1)}_t, \cdots, \alpha^{(B)}_t)^\dagger, 
\;\; \alpha_t^{(k)} > 0 \;\; \forall \; k.
$${#eq-alpha_dir}
We then choose as a prior distribution for the barcode frequency values
$\pi(\underline{f}_t)$ a Dirichlet distribution of the form
$$
\pi(\underline{f}_t) = \frac{
    \Gamma\left(\sum_{b=1}^B\alpha_t^{(b)}\right)
}{
    \prod_{b=1}^B\Gamma(\alpha_t^{(b)})
} 
\prod_{b=1}^B (f_t^{(b)})^{\alpha_t^{(b)}-1},
$${#eq-dirichlet}
where $\Gamma(\cdot)$ is the gamma function. @eq-dirichlet can be much more
compactly written as
$$
\underline{f}_t \sim \operatorname{Dirichlet}(\underline{\alpha}_t),
$${#eq-freq_prior}
where the symbol "$\sim$" is read "distributed as". Again, since the ultimate
calculation will be performed numerically, we can hide all of the mathematical
details in @eq-dirichlet with the simple expression in @eq-freq_prior. We are
free to choose the value of the parameters $\underline{\alpha}_t$ according to
our prior information. A reasonable choice is to assume that before observing
any data at all, we have no reason to believe any barcode is in higher abundance
compared to any other. This might not be entirely true at the beginning of the
experiment, where we expect the neutral barcodes to be in much higher frequency
compared to the mutant barcodes. Nevertheless, we have less certainty about the
following time points, making our "uniform prior over the simplex" a reasonable
choice. Therefore, we define
$$
\alpha_t^{(b)} = 1 \;\; \forall \; b.
$${#eq-alpha_choice}

For the likelihood term $\pi(\underline{r}_t \mid \underline{f}_t)$, we assume a
very simple multinomial model. The story behind this choice goes as follows:
Given a "true" underlying barcode frequency, when the sequencing machine reads a
particular barcode, this is equivalent to having thrown a $B$-sided die, where
the face of the die indicates the barcode that is read. This die is then rolled
$R_t = \sum_b r_t^{(b)}$ times, giving the values of the reads we observe at
time $t$. Notice that this model does not account for noise that might come from
sources such as genetic drift, DNA extraction, and PCR amplification. Although
these sources of noise are important, it is not clear that we have enough
statistical power to make meaningful statements about them given our data (more
on this later on). We can then compactly write the likelihood function as
$$
\underline{r}_t \mid \underline{f}_t \sim
\operatorname{Multinomial}(R_t, \underline{f}_t).
$${#eq-freq_likelihood}

This simple Multinomial-Dirichlet model for the probability of the frequency
values is convenient because it is analytically tractable. Specifically, the
Dirichlet distribution is the so-called *conjugate prior* to the multinomial
distribution. This means the posterior distribution $\pi(\underline{f}_t \mid
\underline{r}_t)$ has a closed-form solution of the same form as the prior
$\pi(\underline{f}_t)$, but with updated parameters. In other words, we can
write (without proof) the posterior distribution as
$$
\underline{f}_t \mid \underline{r}_t \sim
\operatorname{Dirichlet}(\underline{\beta}_t),
$${#eq-freq_posterior}
where the $\underline{\alpha}_t$ parameters are updated to $\underline{\beta}$
after observing the data. These updated parameters are of the form
$$
\beta_t^{(b)} = \alpha_t^{(b)} + r_t^{(b)}.
$$
Simply put, when the likelihood function is multinomial, and the prior 
distribution is Dirichlet, the posterior distribution is also Dirichlet, where
the parameters of the posterior are the parameters of the prior plus the number
of observations of the corresponding output. This fact will become handy
as we build the entire posterior distribution for our inference as defined in
@eq-split_posterior.