### 2023-03-25 doc/_fitness_model.qmd

---
editor:
    render-on-save: true
---

## Preliminaries on mathematical notation
Before jumping directly into what concerns us here, let us establish the
mathematical notation to be used throughout this document. We define (column)
vectors as underlined lowercase symbols such as
$$
\underline{x} = \begin{bmatrix}
    x_1\\
    x_2\\
    \vdots\\
    x_N
\end{bmatrix}.
$${#eq-vec_def}

In the same way, we define matrices as double-underline uppercase symbols such
as $\underline{\underline{A}}$.

## Fitness model {#sec-fitness_model}

Empirically, we have seen that the barcode trajectories follow an exponential
function of the form
$$
f_{t+1}^{(b)} = f_{t}^{(b)} \mathrm{e}^{(s^{(b)} - \bar{s}_t)\tau},
$${#eq-fitness}
where $f_{t}^{(b)}$ is the frequency of barcode $b$ at time $t$, $s^{(b)}$ is
the relative fitness with respect to the ancestral strain---the quantity we want
to infer from the data---$\bar{s}_t$ is the mean fitness of the culture at time
$t$, and $\tau$ is the time pass between time $t$ and $t+1$. It is important to
notice that the frequencies $f_{t}^{(b)}$ are not the allele frequencies in the
population (most of the culture is not sequenced since it is not barcoded), but
rather the inferred frequencies from the total number of reads. A way to 
conceptualize this is that we assume exponential growth in the *number of cells*
$N_t^{(b)}$ of the form
$$
N_{t+1}^{(b)} = N_{t}^{(b)} \mathrm{e}^{(s^{(b)} - \bar{s}_t)\tau},
$${#eq-ncells}
for every barcode $b$. However, when we sequence barcodes, we do not directly
measure the number of cells, but some number of reads $r_t^{(b)}$ that map to
barcode $b$. In the simplest possible scenario, we assume
$$
r_{t}^{(b)} \propto N_{t}^{(b)},
$${#eq-r_to_n}
where, importantly, the proportionality constant depends on the total number of
reads for the time $t$ library, which might vary from library to library.
Therefore, to compare the number of reads between libraries at different time
points, we must somehow normalize them to the same scale. The simplest form is
to define a relative abundance, i.e., a frequency with respect to the total
number of reads,
$$
f_{t}^{(b)} \equiv \frac{r_{t}^{(b)}}{\sum_{b'} r_{t}^{(b')}}.
$${#eq-r_to_n_norm}
This is the frequency @eq-fitness describes.

Our ultimate objective is to infer the relative fitness $s^{(m)}$ for each of
the $M$ mutants in the experiment. To do so, we must take into account the three
main sources of uncertainty in our model:

1. Uncertainty in the determination of frequencies. Our model relates
frequencies between adjacent time points to the fitness of the corresponding
strain. However, we do not measure frequencies directly. Instead, our data
consists of a vector of barcode counts $\underline{r}_t$ for each of the $T$
time points in which the measurements were taken.
2. Uncertainty in the value of the population mean fitness. As we will see later
on, we use neutral lineages barcodes, which, by definition have mean fitness
$s^{(n)}$ to determine the value of the population mean fitness $\bar{s}_t$.
This inference will come with its own uncertainty that must be propagated to the
value of the relative fitness for the mutants.
3. Uncertainty in each of the mutant's fitness values.

To account for all these sources of uncertainty in a principled way, we develop
a Bayesian inference pipeline.