### 2023-03-27 doc/_bayes_mutfit.qmd
finish writing model section

---
editor:
    render-on-save: true
---

### Mutant relative fitness uncertainty $\pi(\underline{s}^M \mid \underline{\bar{s}}_T, \underline{\underline{F}}, \underline{\underline{R}})$ {#seq-bayes_mutfit}

The last piece of our inference is the piece that we care about the most: the
probability distribution of all the mutants' relative fitness, given the
inferred population mean fitness and the frequencies. First, we assume that all
fitness values are independent of each other. This allows us to write
$$
\pi(
    \underline{s}^M \mid 
    \underline{\bar{s}}_T, \underline{\underline{F}}, \underline{\underline{R}}
) = 
\prod_{m=1}^M \pi(
    s^{(m)} \mid
    \underline{\bar{s}}_T, \underline{\underline{F}}, \underline{\underline{R}}
).
$${#eq-mutfit_indep}
Furthermore, as was the case with the population mean fitness, our fitness
model relates frequencies, not raw reads. Moreover, the fitness value of mutant
$m$ only depends on the frequencies of such mutant. Therefore, we can simplify
the conditioning to
$$
\pi(
    s^{(m)} \mid
    \underline{\bar{s}}_T, \underline{\underline{F}}, \underline{\underline{R}}
) = 
\pi(s^{(m)} \mid \underline{\bar{s}}_T, \underline{f}^{(m)}),
$${#eq-mutfit_cond_simple}
where
$$
\underline{f}^{(m)} = (f_0^{(m)}, f_1^{(m)}, \ldots, f_T^{(m)})^\dagger,
$${#eq-mutfit_timeseries_def}
is the vector containing the frequency time series for mutant $m$. Writing 
Bayes' theorem for the right-hand side of @eq-mutfit_cond_simple results in
$$
\pi(s^{(m)} \mid \underline{\bar{s}}_T, \underline{f}^{(m)}) \propto
\pi(\underline{f}^{(m)} \mid \underline{\bar{s}}_T, s^{(m)})
\pi(s^{(m)} \mid \underline{\bar{s}}_T).
$${#eq-mutfit_bayes}
Notice the conditioning on the mean fitness values $\underline{\bar{s}}_T$ is 
not inverted since we already inferred these values.

Following the logic used in @sec-bayes_meanfit, let us define
$$
\underline{\gamma}^{(m)} = 
(\gamma_0^{(m)}, \gamma_1^{(m)}, \ldots, \gamma_{T-1}^{m})^\dagger,
$${#eq-gamma_vec_def}
where each entry $\gamma_t^{(m)}$ is defined by @eq-gamma_def. In the same way
we rewrote the joint distribution between two adjacent time point frequencies 
to the joint distribution between one of the frequencies and the ratio of both 
frequencies in @eq-joint_freq_gamma, we can rewrite the joint distribution of
the frequency time series for mutant $m$ as
$$
\pi(\underline{f}^{(m)} \mid \underline{\bar{s}}_T, s^{(m)}) =
\pi(f_0^{(m)}, \underline{\gamma}^{(m)} \mid \underline{\bar{s}}_T, s^{(m)}).
$${#eq-joint_freq_gammas_mutfit}
One can think about @eq-joint_freq_gammas_mutfit as saying that knowing the
individual frequencies at each time point contain equivalent information as
knowing the initial frequency and the subsequent ratios of frequencies. This is
because if we want to know the value of $f_1^{(m)}$ given the ratios, we only
need to compute
$$
f_1^{(m)} = \gamma_0^{(m)} f_0^{(m)}.
$${#eq-f1_from_ratio}
Moreover, if we want to know $f_2^{(m)}$, we have
$$
f_2^{(m)} = \gamma_1^{(m)} f_1^{(m)} =
\gamma_1^{(m)} \left(\gamma_0^{(m)} f_0^{(m)}\right),
$${#eq-f2_from_ratio}
and so on. We can then write the joint distribution on the right-hand side of
@eq-joint_freq_gammas_mutfit as a product of conditional distributions of the
form
$$
\begin{aligned}
\pi(f_0^{(m)}, \underline{\gamma}^{(m)} \mid \underline{\bar{s}}_T, s^{(m)}) =
&\pi(
    f_0^{(m)} \mid 
    \gamma_0^{(m)}, \ldots, \gamma_{T-1}^{(m)}, \underline{\bar{s}}_T, s^{(m)}
) \times \\
&\pi(
    \gamma_0^{(m)} \mid 
    \gamma_1^{(m)}, \ldots, \gamma_{T-1}^{(m)}, \underline{\bar{s}}_T, s^{(m)}
) \times \\
&\pi(
    \gamma_1^{(m)} \mid 
    \gamma_2^{(m)}, \ldots, \gamma_{T-1}^{(m)}, \underline{\bar{s}}_T, s^{(m)}
) \times \\
&\vdots \\
&\pi(
    \gamma_{T-2}^{(m)} \mid \gamma_{T-1}^{(m)}, \underline{\bar{s}}_T, s^{(m)}
) \times \\
&\pi(\gamma_{T-1}^{(m)} \mid \underline{\bar{s}}_T, s^{(m)}).
\end{aligned}
$${#eq-mutfit_joint_to_cond}
Writing the fitness model in @eq-fitness as
$$
\gamma_t^{(m)} = \frac{f_{t+1}^{(m)}}{f_t^{(m)}} = 
\mathrm{e}^{(s^{(m)} - s_t)\tau},
$$
reveals that the value of each of the ratios $\gamma_t^{(m)}$ only depends on
the corresponding fitness value $\bar{s}_t$ and the relative fitness $s^{(m)}$.
Therefore, we can remove most of the conditioning on the right-hand side of 
@eq-mutfit_joint_to_cond, resulting in a much simpler joint distribution of the
form
$$
\begin{aligned}
\pi(f_0^{(m)}, \underline{\gamma}^{(m)} \mid \underline{\bar{s}}_T, s^{(m)}) =
&\pi(f_0^{(m)} \mid \gamma_0^{(m)}) \times \\
&\pi(\gamma_0^{(m)} \mid \bar{s}_0, s^{(m)}) \times \\
&\pi(\gamma_1^{(m)} \mid \bar{s}_1, s^{(m)}) \times \\
&\vdots \\
&\pi(\gamma_{T-2}^{(m)} \mid \bar{s}_{T-2}, s^{(m)}) \times \\
&\pi(\gamma_{T-1}^{(m)} \mid \bar{s}_{T-1}, s^{(m)}),
\end{aligned}
$${#eq-mutfit_joint_to_cond_simple}
where for the first term on the right-hand side of
@eq-mutfit_joint_to_cond_simple we apply the same logic as in
@eq-freq_cond_gamma to remove all other dependencies. We emphasize that although
@eq-mutfit_joint_to_cond_simple looks like a series of independent inferences,
the value of the relative fitness $s^{(m)}$ is shared among all of them. This
means that the parameter is not inferred individually for each time point,
resulting in different estimates of the parameter, but each time point
contributes independently to the inference of a single estimate of $s^{(m)}$.

Using equivalent arguments to those in @sec-bayes_meanfit, we assume
$$
f_0^{(m)} \mid \gamma_0^{(m)} \sim 
\operatorname{Uniform}\left(0, \frac{1}{\gamma_0^{(m)}} \right),
$$
and
$$
\gamma_t^{(m)} \mid \bar{s}_t, s^{(m)}, \sigma^{(m)} \sim 
\log\mathcal{N}\left(s^{(m)} - \bar{s}_t, \sigma^{(m)} \right),
$${#eq-mutfit_lognormal}
where we add the nuisance parameter $\sigma^{(m)}$ to the inference. Notice that
this parameter is not indexed by time. This means that we assume the deviations
from the theoretical prediction do not depend on time, but only on the mutant.
Adding the nuisance parameter demands us to update @eq-mutfit_bayes to
$$
\pi(
    s^{(m)}, \sigma^{(m)} \mid \underline{\bar{s}}_T, \underline{f}^{(m)}
) \propto
\pi(\underline{f}^{(m)} \mid \underline{\bar{s}}_T, s^{(m)}, \sigma^{(m)})
\pi(s^{(m)}) \pi(\sigma^{(m)}),
$${#eq-mutfit_bayes_full}
where we assume independent priors for both parameters. We also removed the
conditioning on the values of the mean fitness as knowing such values does not
change our prior information about the possible range of values that the
parameters can take. As with the priors on @sec-bayes_meanfit, we will assign
weakly-informative priors to these parameters.

#### Summary

With all pieces in place, we write the full inference of the relative fitness
values as
$$
\pi(
    \underline{s}^M ,\underline{\sigma}^M \mid 
    \underline{\bar{s}}_T, \underline{\underline{F}}
) \propto
\prod_{m=1}^M \left\{ 
    \pi(f_0^{(m)} \mid \gamma_0^{(m)})
    \prod_{t=0}^{T-1} \left[
        \pi(\gamma_t^{(m)} \mid \bar{s}_t, s^{(m)}, \sigma^{(m)})
    \right]
    \pi(s^{(m)}) \pi(\sigma^{(m)})
\right\},
$${#eq-mutfit_full_inference}
where
$$
f_0^{(m)} \mid \gamma_0^{(m)} \sim 
\operatorname{Uniform}\left(0, \frac{1}{\gamma_0^{(m)}} \right),
$${#eq-mutfit_eq1}
$$
\gamma_t^{(m)} \mid \bar{s}_t, s^{(m)}, \sigma^{(m)} \sim 
\log\mathcal{N}\left(s^{(m)} - \bar{s}_t, \sigma^{(m)} \right),
$${#eq-mutfit_eq2}
$$
s^{(m)} \sim \mathcal{N}(0, \sigma_{s^{(m)}}),
$${#eq-mutfit_eq3}
and
$$
\sigma^{(m)} \sim \operatorname{Half}-\mathcal{N}(0, \sigma_{\sigma^{(m)}}),
$${#eq-mutfit_eq4}
where $\sigma_{s^{(m)}}$ and $\sigma_{\sigma^{(m)}}$ are free parameters we can
choose.