### 2023-03-30 doc/_mcmc_sampling.qmd

---
editor:
    render-on-save: true
---
## Markov Chain Monte Carlo sampling of posterior {#sec-mcmc_sampling}