### 2023-04-12 doc/_mcmc_sampling.qmd
keep working on _mcmc_sampling section

---
editor:
    render-on-save: true
---
## Markov Chain Monte Carlo sampling of posterior {#sec-mcmc_sampling}

The model proposed in @sec-bayesian_inference presents a daunting computational
task as it demands the simultaneous inference of > $10^4$ parameters. Sampling
such high-dimensional posterior distribution can be extremely slow, even for
state-of-the-art Markov Chain Monte Carlo samplers. Nevertheless, given the
structure of our problem, we can get around this issue by focusing on the
marginal individual mutants' relative fitness posterior distributions. This is
perfectly justified as we are only interested in these marginal distributions
$\pi(s^{(m)} \mid \underline{\underline{R}})$, rather than the full joint
posterior distribution.

To make the inference task computationally feasible, we divide the process into
two steps:

1. Sample out of the posterior distribution of the population mean fitness using
the information of the neutral lineages only.
2. Sample out of each mutant's relative fitness posterior distribution given the
knowledge acquired in step 1.

In what follows, we describe the modification of the model presented in
@eq-mutfit_full_inference to make Markov Chain Monte Carlo sampling of the
posterior distribution computationally feasible.

### Sample population mean fitness posterior distribution

The full joint posterior distribution for our experiment $\pi(\underline{s}^M,
\underline{\sigma}^M, \underline{\bar{s}}_T, \underline{\sigma}_T,
\underline{\underline{F}} \mid \underline{\underline{R}})$ demands the
simultaneous inference of all of the frequencies $\underline{\underline{F}}$,
the population mean fitness values for every pair of time points
$\underline{\bar{s}}_T$, the mutants relative fitness $\underline{s}^M$, and the
nuisance parameters inherited from the likelihood functions
$\underline{\sigma}^M$ and $\underline{\sigma}_T$. Throughout
@sec-bayesian_inference, we describe a series of independence assumptions that
enormously simplified the problem. Nevertheless, these assumptions do not remove
all of the computational complexity, as we still need to infer all of the
frequency values for each of the $~ 10^3-10^4$ unique barcodes.

The first step towards beating the full posterior distribution to a
computationally tractable form is to focus on the inference of the population
mean fitness, ignoring the mutants' relative fitness values. This is
mathematically sound, given the structure of our problem, as described in
@sec-bayesian_inference. More specifically, @eq-split_posterior shows that we
can split the posterior into the product of three terms, where the term
involving the mean fitness values $\underline{\bar{s}}_T$ do not depend on the
mutants relative fitness. Thus, we are left with the challenge of sampling from
a "simpler" posterior distribution with less parameters,
$$
\pi(
    \underline{\bar{s}}_T, \underline{\sigma}_T, \underline{\underline{F}} \mid
    \underline{\underline{R}}
) = 
\pi(\underline{\bar{s}}_T, \underline{\sigma}_T \mid \underline{\underline{F}})
\pi(\underline{\underline{F}} \mid \underline{\underline{R}}),
$${#eq-simpler_meanfit_posterior}
where the right-hand side splits the joint distribution into the product of
conditional distributions, as in @eq-split_posterior.
@eq-simpler_meanfit_posterior already removed ~$10^3$ parameters from the 
inference problem, but we still have the simultaneous inference of all 
frequencies $\underline{\underline{F}}$. To make further progress, we can follow
the independence assumptions described in @sec-bayes_meanfit, which led to
@eq-meanfit_indep, to write the posterior distribution as a product of the form
$$
\pi(\underline{\bar{s}}_T, \underline{\sigma}_T \mid \underline{\underline{F}})
\pi(\underline{\underline{F}} \mid \underline{\underline{R}}) 
= \prod_{t=1}^{T-1} 
\pi(\bar{s}_t, \sigma_t, \mid \underline{f}_t, \underline{f}_{t+1})
\pi(\underline{f}_t \mid \underline{r}_t)
\pi(\underline{f}_{t+1} \mid \underline{r}_{t+1}).
$${#eq-meanfit_posterior_indep}
If we sample from the posterior distribution of one of the $T-1$ terms on the
right-hand side of @eq-meanfit_posterior_indep, we again enormously reduce the
number of parameters that need to be inferred simultaneously, as we only focus
on the frequencies $\underline{f}_t$, and $\underline{f}_{t+1}$, relevant for a
single mean fitness value $\bar{s}_t$ rather than all frequencies
$\underline{\underline{F}}$. Nevertheless, the large number of unique barcodes
makes this still a challenging inference problem.

The frequencies $\underline{f}_t$, and $\underline{f}_{t+1}$, and the standard
deviation $\sigma_t$ represent nuisance parameters we ignore at the end of the
inference procedure. In other words, what we care about is the marginal
distribution of the mean fitness given the data, i.e.,
$$
\pi(\bar{s}_t \mid \underline{r}_t, \underline{r}_{t+1}) =
\int d\sigma_t\;
\int d^B \underline{f}_t\;
\int d^B \underline{f}_{t+1}\;
\pi(\bar{s}_t, \sigma_t, \mid \underline{f}_t, \underline{f}_{t+1})
\pi(\underline{f}_t \mid \underline{r}_t)
\pi(\underline{f}_{t+1} \mid \underline{r}_{t+1}),
$${#eq-meanfit_posterior_marginal}
where we remind the reader that $B$ is the total number of unique barcodes. 
These integrals with respect to all barcode frequencies pose the ultimate
computational challenge we must overcome. To this goal, notice that the
inference of the mean fitness $\bar{s}_t$ makes use only of the neutral lineages
frequencies. Thus, it is convenient to split the $\underline{f}_t$ vectors into
the neutral and mutant-only frequency arrays and write
$$
\pi(\bar{s}_t, \sigma_t, \mid \underline{f}_t, \underline{f}_{t+1})
\pi(\underline{f}_t \mid \underline{r}_t)
\pi(\underline{f}_{t+1} \mid \underline{r}_{t+1}) = 
\pi(
    \bar{s}_t, \sigma_t, \mid 
    \underline{f}_t^N, \underline{f}_{t+1}^N
)
\pi(
    \underline{f}_t^N, \underline{f}_t^M \mid \underline{r}_t
)
\pi(
    \underline{f}_{t+1}^N, \underline{f}_{t+1}^M \mid \underline{r}_{t+1}
),
$${#eq-meanfit_split_f}
where the first term on the right-hand side of @eq-meanfit_split_f only includes
the neutral frequencies needed to infer the population mean fitness. The
integral in @eq-meanfit_posterior_marginal can then be split for neutral and
mutant lineages, resulting in a rather convoluted equation of the form 
$$
\begin{aligned}
\pi(\bar{s}_t \mid \underline{r}_t, \underline{r}_{t+1}) =
&\int d\sigma_t\;
\int d^N \underline{f}_t^N \;
\int d^N \underline{f}_{t+1}^N 
\int d^M \underline{f}_t^M \;
\int d^M \underline{f}_{t+1}^M \times \\
&\pi(
    \bar{s}_t, \sigma_t, \mid 
    \underline{f}_t^N, \underline{f}_{t+1}^N
)
\pi(
    \underline{f}_t^N, \underline{f}_t^M \mid \underline{r}_t
)
\pi(
    \underline{f}_{t+1}^N, \underline{f}_{t+1}^M \mid \underline{r}_{t+1}
),
\end{aligned}
$${#eq-meanfit_split_f_integral}
where, again, we remind the reader that $N$ is the number of neutral barcodes
and $M$ the number of mutant barcodes. Despite the convoluted form, we can
rewrite @eq-meanfit_split_f_integral to get clarity on how to reduce the
dimensionality of the problem even further. Distributing terms in
@eq-meanfit_split_f_integral results in
$$
\begin{aligned}
\pi(\bar{s}_t \mid \underline{r}_t, \underline{r}_{t+1}) =
&\int d\sigma_t\;
\int d^N \underline{f}_t^N \;
\int d^N \underline{f}_{t+1}^N 
\pi(
    \bar{s}_t, \sigma_t, \mid 
    \underline{f}_t^N, \underline{f}_{t+1}^N
) \times \\
&\int d^M \underline{f}_t^M \;
\pi(
    \underline{f}_t^N, \underline{f}_t^M \mid \underline{r}_t
) \times \\
&\int d^M \underline{f}_{t+1}^M \;
\pi(
    \underline{f}_{t+1}^N, \underline{f}_{t+1}^M \mid \underline{r}_{t+1}
).
\end{aligned}
$$