### 2023-07-20 doc/_bayesian_inference.qmd
making progress on writing the paper

---
editor:
    render-on-save: true
bibliography: references.bib
csl: ieee.csl
---
## Bayesian inference {#sec-bayesian_inference}

As defined in @sec-fitness_model, our ultimate objective is to infer the vector
of relative fitness values
$$
\underline{s}^M = (s^{(1)}, s^{(2)}, \ldots, s^{(M)})^\dagger,
$${#eq-fitness_mut_vec}
where $^\dagger$ indicates the transpose. Our data consists of an $T \times B$
matrix $\underline{\underline{R}}$, where $B = M + N$ is the number of unique
barcodes given by the sum of the number of unique mutant barcodes $M$ and the
number of unique neutral barcodes $N$, and $T$ is the number of time points
where measurements were taken. The data matrix is then of the form
$$
\underline{\underline{R}} = \begin{bmatrix}
- & \underline{r}_1 & - \\
- & \underline{r}_2 & - \\
 & \vdots & \\
- & \underline{r}_T & - \\
\end{bmatrix},
$${#eq-R_mat}
where each row $\underline{r}_t$ is a $B$-dimensional array containing the
raw barcode counts at time $t$. We can further split each vector
$\underline{r}_t$ into two vectors of the form
$$
\underline{r}_t = \begin{bmatrix}
\underline{r}_t^{N} \;
\underline{r}_t^{M}
\end{bmatrix},
$${#eq-r-vec_split}
i.e., the vector containing the neutral lineages barcode counts
$\underline{r}_t^{N}$ and the corresponding vector containing the mutant barcode
counts $\underline{r}_t^{M}$. Our objective is to compute the joint probability
distribution of all relative fitness values given our data. We can express this 
joint distribution using Bayes theorem as
$$
\pi(\underline{s}^M \mid \underline{\underline{R}}) = \frac{
\pi(\underline{\underline{R}} \mid \underline{s}^M) 
\pi(\underline{s}^M)}
{\pi(\underline{\underline{R}})},
$${#eq-bayes_obj}
where hereafter $\pi(\cdot)$ defines a probability density function unless
otherwise stated. We can ignore the denominator on the right-hand side of
@eq-bayes_obj, since it serves as a normalization constant that does not depend
on our parameter of interest, simply writing
$$
\pi(\underline{s}^M \mid \underline{\underline{R}}) \propto
\pi(\underline{\underline{R}} \mid \underline{s}^M) 
\pi(\underline{s}^M)
$${#eq-bayes_obj_propto}

Although @eq-bayes_obj and @eq-bayes_obj_propto seem simple enough, recall that
@eq-fitness relates barcode frequency values and the population mean fitness to
the mutant relative fitness. Therefore, we must include these nuisance
parameters as part of our inference problem. We can always marginalize the joint
distribution to obtain the distribution we are interested in. To include these
parameters, let 
$$
\underline{\bar{s}}_T = (\bar{s}_1, \bar{s}_2, \ldots, \bar{s}_{T-1})^\dagger,
$${#eq-pop_fitness_vec}
be the vector containing the $T-1$ population mean fitness we compute from the
$T$ time points where measurements were taken. We have $T-1$ since the value
of any $\bar{s}_t$ requires time points $t$ and $t+1$. Furthermore, let the
matrix $\underline{\underline{F}}$ be a $T \times B$ matrix containing all
frequency values. As with @eq-R_mat, each row of $\underline{\underline{F}}$
can be split into two components of the form
$$
\underline{f}_t = \begin{bmatrix}
\underline{f}_t^M \;
\underline{f}_t^N
\end{bmatrix},
$${#eq-f_vec}
to separate the corresponding mutant and neutral barcode frequencies. With these
variables in hand, the full inference problem we must solve takes the form
$$
\pi(
    \underline{s}^M, \underline{\bar{s}}_T, \underline{\underline{F}} \mid
    \underline{\underline{R}}
) \propto
\pi(
    \underline{\underline{R}} \mid
    \underline{s}^M, \underline{\bar{s}}_T, \underline{\underline{F}}
)
\pi(
    \underline{s}^M, \underline{\bar{s}}_T, \underline{\underline{F}}
).
$${#eq-bayes_full}
We can numerically integrate out all nuisance parameters to recover the marginal
distribution we care about, i.e.,
$$
\pi(\underline{s}^M \mid \underline{\underline{R}}) =
\int d^{T-1}\underline{\bar{s}}_T
\int d^{B}\underline{f}_1 \cdots
\int d^{B}\underline{f}_T
\;
\pi(
    \underline{s}^M, \underline{\bar{s}}_T, \underline{\underline{F}} \mid
    \underline{\underline{R}}
).
$${#eq-bayes_full_marginal}

### Factorizing the posterior distribution {#seq-split_posterior}

The left-hand side of @eq-bayes_full is extremely difficult to work with.
However, we can take advantage of the structure of our inference problem to
rewrite it in a more manageable form. Specifically, we rewrite the joint
probability distribution as a product of conditional distributions of the form
$$
\pi(
    \underline{s}^M, \underline{\bar{s}}_T, \underline{\underline{F}} \mid
    \underline{\underline{R}}
) =
\pi(
    \underline{s}^M \mid \underline{\bar{s}}_T, \underline{\underline{F}},
    \underline{\underline{R}}
)
\pi(
    \underline{\bar{s}}_T \mid \underline{\underline{F}}, 
    \underline{\underline{R}}
)
\pi(\underline{\underline{F}} \mid \underline{\underline{R}}).
$${#eq-split_posterior}
Written in this form, @eq-split_posterior captures the three sources of 
uncertainty listed in @sec-fitness_model in each of the terms. Starting from
right to left, the first term on the right-hand side of @eq-split_posterior
accounts for the uncertainty when inferring the frequency values given the
barcode reads. The second term accounts for the uncertainty in the values of
the mean population fitness at different time points. The last term accounts for
the uncertainty in the parameter we care about---the mutants' relative
fitnesses. We invite the reader to check the supplementary material where we
work through each of these terms to build the posterior distribution to sample
from.