### 2023-07-20 doc/_intro.qmd

---
editor:
    render-on-save: true
bibliography: references.bib
csl: ieee.csl
---

The advent of DNA barcoding and high-throughput sequencing has opened new venues
to understand microbial evolutionary dynamics with an unprecedented level of
temporal resolution [@levy2015; @nguyenba2019a; @ascensao2023]. These
experimental efforts rely on our ability to reliably infer the relative fitness
of an ensemble of diverse genotypes. Moreover, inferring these fitness values
over an ensemble of environmental conditions can help us determine the
phenotypic diversity of a rapid adaptation process @kinsler2020.

As with any other sequencing-based quantification, tracking lineages via DNA
barcode sequencing is inexorably accompanied by noise sources coming from
experimental manipulation of the microbial cultures, DNA extraction and
sequencing library preparation that involves multiple rounds of PCR
amplification, and the sequencing process itself. Thus, accounting for the
uncertainty when inferring the parameters we care about from the data is a key
step to drawing reliable conclusions. Bayesian statistics presents a paradigm
with which to account for all known sources of uncertainty in a principled way
@eddy2004a. This, combined with the development of modern Markov Chain Monte
Carlo sampling algorithms have boosted a resurgence of Bayesian methods in
different fields @efron2013a.

We present a Bayesian inference pipeline to properly quantify our uncertainty
about the parametric information we can extract from high-throughput competitive
fitness assays. In these assays, the fitness of an ensemble of genotypes is
determined relative to a reference genotype [@kinsler2020; @ascensao2023] [`add
reference to fig`]. Small changes to our model allow us to account for relevant
experimental variations of these competition assays: determination of multiple
fitness values per genotype for growth-dilution cycles on different
environments, and a hierarchical model to account for information from
independent experimental replicates. We benchmark the ability of our pipeline to
infer fitness parameters against synthetic data with known ground truth and
re-analyze published data from @kinsler2020 to assess the method's performance
on real experimental measurements. The method is accompanied with a `Julia`
package to readily implement the inference pipeline with state-of-the-art
scientific computing software. [`here mention something about the Python wrapper
if appropriate`].