### 2023-07-26 doc/_experiment.qmd
continue to make progress on main results

---
editor:
    render-on-save: true
bibliography: references.bib
csl: ieee.csl
---

## Experimental setup

Our inference pipeline is designed for the type of competition assays presented
in @kinsler2020. In these experiments, an ensemble of genotypes is pooled
together with an unlabeled reference strain that, initially, represents the vast
majority ($\geq 90\%$) of the cells in the culture (see schematic in
@fig-01(A)). Furthermore, a subfraction of labeled genotypes equivalent to
the unlabeled reference strain---hereafter defined as *neutral* lineages---are
spiked in at a relatively high abundance ($\approx 3-5\%$), the rest of the
culture being left for the ensemble of genotypes of interest.

To determine the relative fitness of the ensemble of genotypes, a series of
growth-dilution cycles are performed on either a single or multiple
environments. In other words, the cultures are grown for some time to then
dilute the culture back into fresh media and repeat for roughly 4-6 cycles,
depending on the initial abundances of the mutants, and their relative growth
rates. The DNA barcodes are sequenced at the end of each growth cycle to
quantify the relative abundance of each of the barcodes. We point the reader to
@kinsler2020 for specific details on these assays for *S. cerevisiae*.
@fig-01(B) presents a typical barcode trajectory where the darker barcode
trajectories represent the neutral lineages.

![**Typical competitive fitness experiment**. (A) Schematic of the typical
experimental design to determine the competitive fitness of an ensemble of
barcoded genotypes. Genotypes are pooled together and grown over multiple
growth-dilution cycles. At the end of each cycle, a sample is processed to
generate a library for amplicon sequencing. (B) Typical barcode trajectory
dataset. From each time point, the relative frequency of each barcode is
determined from the total number of reads. Different shades of blue represent
different relative fitness. Darker gray lines define the typical trajectory of
neutral lineages.](./figs/fig01){#fig-01}