### 2023-07-26 doc/_single_results.qmd

---
editor:
    render-on-save: true
bibliography: references.bib
csl: ieee.csl
---

### Inference on a single dataset

To assess the inference pipeline performance we applied it to a simulated
dataset with known ground truth (See `[ref to SI]` for details on simulation).
@fig-02(A) shows the structure of the synthetic dataset. The majority of
barcodes of interest (dim color lines) are adaptive compared to the neutral
barcodes ($s^{(m)} > 0$). This experimental condition is required to reliably
measure the relative fitness since all barcodes of interest occupy a small
fraction of the population at the beginning of the experiment `[ref SI section
on detrimental mutants]`. Although the barcode frequency trajectories look
relatively smooth, our fitness model requires the computation of the log
frequency ratio between adjacent time points. @fig-02(B) shows such data
transformation. Here, we can better appreciate the variability in the input that
goes into our model. This is especially clear for the darker lines representing
the neutral barcodes since they are supposed to be identically distributed under
the log frequency ratio transformation.

A way to visualize the performance of our inference pipeline in fitting our
fitness model to the observed data is to compute the so-called posterior
predictive checks (PPC). In short, the PPC consists of bootstrapping the
generation of synthetic datasets in agreement with the results from the
inference results. In other words, we use the resulting parameter values from
the MCMC chain to generate possible datasets in agreement with the inferred
values. @fig-02(C) shows these results for all neutral lineages (upper left
corner plot) and a few representative non–neutral barcodes. The dark color line
shows the median value we would expect the data to be found, while the shaded
regions show the 68\% and 95\% credible region, i.e. the regions where we expect
to find the data with the corresponding probability---or in terms of our
parameter, the $X\%$ credible region is the interval where we expect the true
parameter to lie with $X\%$ probability. The main advantage of our method is the
natural interpretability of these credible regions. A common mistake in the
literature is to confuse frequentist confidence intervals with Bayesian credible
regions when they are not equivalent @morey2016.

To capture the global performance of the model, @fig-02(D) compares the known
groud truth with the inferred relative fitness value for all barcodes of
interest. There is a great degree of correspondence between these values, with
the error bars, representing the 68\% credible region for the parameter value
crossing the identity line for most barcodes. This latter point is made clear
with @fig-02(E) where the empirical cumulative distribution function (ECDF) of
the absolute differences between the median inferred value and the true value
shows that all median inferred values are within 0.3 units of relative fitness
from the ground truth. Furthermore, because of the Bayesian scope of our method,
we do not need to limit ourselves to a point estimate of this quantity, but
report (asymmetric) error bars that fully quantify our degree of belief in the
inferred value as shown on top of the panels in @fig-02(C).

![**Single dataset inference**. (A) Frequency trajectories that represent the
raw data going into the inference. (B) Log frequency ratio between two adjacent
time points used by the inference pipeline. Darker lines represent the neutral
barcodes. Notice that these transformed data are much more noisy than the
seemingly smooth frequency trajectories. (B) Examples of the posterior
predictive checks for all neutral lineages (upper left panel) and a subset of
representative mutant lineages. Asymmetric error bars represent the 68\%
credible region on the mutant relative fitness marginal distribution. (C)
Comparison between the ground truth fitness value from the logistic-growth
simulation and the inferred fitness value. Gray error bars represent the 68\%
posterior credible region. (D) The empirical cumulative distribution function of
the absolute inference between the ground truth fitness value and the median of
the marginal distributions for each mutant. This plot reveals that all median
inferred values are within 0.3 units of relative fitness from the ground
truth.](./figs/fig02){#fig-02}