### 2023-09-06 doc/_fitness_model.qmd

---
editor:
    render-on-save: true
bibliography: references.bib
csl: ieee.csl
---

## Preliminaries on mathematical notation
Before jumping directly into the Bayesian inference pipeline, let us establish
the mathematical notation used throughout this paper. We define
(column) vectors as underlined lowercase symbols such as
$$
\underline{x} = \begin{bmatrix}
    x_1\\
    x_2\\
    \vdots\\
    x_N
\end{bmatrix}.
$${#eq-vec_def}
In the same way, we define matrices as double-underline uppercase symbols such
as
$$
\underline{\underline{A}} =
\begin{bmatrix}
    A_{11} & A_{12} & \cdots & A_{1N}\\
    A_{21} & A_{22} & \cdots & A_{2N}\\
    \vdots & \vdots & \ddots & \vdots\\
    A_{M1} & A_{M2} & \cdots & A_{MN}\\
\end{bmatrix}.
$${#eq-mat_def}

## Fitness model {#sec-fitness_model}

Empirically, the barcode relative frequency trajectories follow an exponential
function of the form
$$
f_{t+1}^{(b)} = f_{t}^{(b)} \mathrm{e}^{(s^{(b)} - \bar{s}_t)\tau},
$${#eq-fitness}
where $f_{t}^{(b)}$ is the frequency of barcode $b$ at time $t$, $s^{(b)}$ is
the relative fitness with respect to the reference strain---the quantity we want
to infer from the data---$\bar{s}_t$ is the mean fitness of the culture at time
$t$, and $\tau$ is the time pass between time $t$ and $t+1$. We can rewrite
@eq-fitness as
$$
\frac{1}{\tau}\ln \frac{f_{t+1}^{(b)}}{f_{t}^{(b)}} = (s^{(b)} - \bar{s}_t).
$${#eq-logfreq}
@eq-logfreq separates the observables---the barcode frequencies---from the
latent parameters we want to infer from the data---the population mean fitness
and the barcode relative fitness. This is ultimately the functional form used in
our inference pipeline. Therefore, the relative fitness is computed by knowing
the log frequency ratio of each barcode throughout the growth-dilution cycles.

It is important to notice that the frequencies $f_{t}^{(b)}$ are not the allele
frequencies in the population (most of the culture is not sequenced since the
reference strain is not barcoded), but rather the relative frequencies in the
total number of sequencing reads. A way to conceptualize this subtle but
important point is to assume exponential growth in the *number of cells*
$N_t^{(b)}$ of the form
$$
N_{t+1}^{(b)} = N_{t}^{(b)} \mathrm{e}^{\lambda^{(b)}\tau},
$${#eq-ncells}
for every barcode $b$ with growth rate $\lambda^{(b)}$. However, when we
sequence barcodes, we do not directly measure the number of cells, but some
number of reads $r_t^{(b)}$ that map to barcode $b$. In the simplest possible
scenario, we assume
$$
r_{t}^{(b)} \propto N_{t}^{(b)},
$${#eq-r_to_n}
where, importantly, the proportionality constant depends on the total number of
reads for the time $t$ library, which might vary from library to library.
Therefore, to compare the number of reads between libraries at different time
points, we must normalize the number of reads to the same scale. The simplest
form is to define a relative abundance, i.e., a frequency with respect to the
total number of reads,
$$
f_{t}^{(b)} \equiv \frac{r_{t}^{(b)}}{\sum_{b'} r_{t}^{(b')}}.
$${#eq-r_to_n_norm}
This is the frequency @eq-fitness describes.

Our ultimate objective is to infer the relative fitness $s^{(m)}$ for each of
the $M$ relevant barcodes in the experiment. To do so, we account for the three
primary sources of uncertainty in our model:

1. Uncertainty in the determination of frequencies. Our model relates
frequencies between adjacent time points to the fitness of the corresponding
strain. However, we do not directly measure frequencies. Instead, our data for
each barcode consists of a length $T$ vector of counts $\underline{r}^{(b)}$ for
each of the $T$ time points in which the measurements were taken.
2. Uncertainty in the value of the population mean fitness. We define neutral
lineages to have fitness $s^{(n)} = 0$, helping us anchor the value of the
population mean fitness $\bar{s}_t$ for each pair of adjacent time points. 
Moreover, we take this parameter as an empirical observable from the data, 
meaning that we do not have a functional form that relates $\bar{s}_t$ to
$\bar{s}_{t+1}$. Thus, we must infer the $T-1$ values of this population mean 
fitness with their uncertainty that must be propagated to the value of
the mutants' relative fitness.
3. Uncertainty in each of the mutants' fitness values.

To account for all these sources of uncertainty in a principled way, in the next
section, we develop a Bayesian inference pipeline.