### 2023-09-06 doc/_hierarchical_results.qmd

---
editor:
    render-on-save: true
bibliography: references.bib
csl: ieee.csl
---

### Accounting for experimental replicates via hierarchical models

Our inference pipeline can be extended to account for multiple experimental
replicates via Bayesian hierarchical models [@betancourt2013]. Briefly, when
accounting for multiple repeated measurements of the same phenomena, there are
two extreme cases one can use to perform the data analysis: On the one hand, we
can treat each measurement as completely independent, losing the power to
utilize multiple measurements when trying to learn a single parameter. This can
hurt the inference since, in principle, the value of our parameter of interest
should not depend on the particular experimental replicate in question. However,
this approach does not allow us to properly "combine" the uncertainties in both
experiments when performing the inference. On the other hand, we can pool all
data together and treat our different experiments as a single measurement with
higher coverage. This loses the subtle differences that come from biotic and
abiotic batch effects, effectively halving the data that goes into our inference
problem.

Hierarchical models present a middle ground between these extremes. First,
hierarchical models rely on the definition of so-called *hyper-parameters*, that
capture the ultimate parameter we are interested in---in our case we have a
hyper-fitness value $\theta^{(m)}$ for each mutant. Second, each experiment
draws randomly from the distribution of this hyper-parameter, allowing for
subtle variability between experiments to be accounted for---for our inference
pipeline, each experimental replicate gets assigned a fitness value $s^{(m,i)}$,
where the extra index indicates the $i$-th experimental replicate.

![**Hierarchical model on experimental replicates**.](./figs/fig03){#fig-03}