### 2023-09-08 doc/_hierarchical_replicates_results.qmd
rename section

---
editor:
    render-on-save: true
bibliography: references.bib
csl: ieee.csl
---

## Accounting for experimental replicates via hierarchical models

Our inference pipeline can be extended to account for multiple experimental
replicates via Bayesian hierarchical models [@betancourt2013]. Briefly, when
accounting for multiple repeated measurements of the same phenomena, there are
two extreme cases one can use to perform the data analysis: On the one hand, we
can treat each measurement as completely independent, losing the power to
utilize multiple measurements when trying to learn a single parameter. This can
hurt the inference since, in principle, the value of our parameter of interest
should not depend on the particular experimental replicate in question. However,
this approach does not allow us to properly "combine" the uncertainties in both
experiments when performing the inference. On the other hand, we can pool all
data together and treat our different experiments as a single measurement with
higher coverage. This loses the subtle differences that come from biotic and
abiotic batch effects, effectively halving the data that goes into our inference
problem.

Hierarchical models present a middle ground between these extremes. First,
hierarchical models rely on the definition of so-called *hyper-parameters*, that
capture the ultimate parameter we are interested in---for our case we have a
hyper-fitness value $\theta^{(m)}$ for each mutant. Second, each experiment
draws randomly from the distribution of this hyper-parameter, allowing for
subtle variability between experiments to be accounted for---for our inference
pipeline, each experimental replicate gets assigned a fitness value $s^{(m,i)}$,
where the extra index indicates the $i$-th experimental replicate. As
schematized by the Bayesian network on @fig-03(A) and further detailed in
`[point to SI section]`, our statistical model can easily be extended to a
hierarchical structure, where the fitness of each relevant barcode on each
experimental replicate is a sample from a global hyper-fitness parameter.

To test the performance of this model, we simulated two experimental replicates
with 1000 unique barcodes (see @fig-03(B-C)) where we randomly sampled a ground
truth hyper-fitness value $\theta^{(m)}$ for each barcode. To capture
experimental batch effects, we sampled a variation from this hyper-fitness value
for each experimental replicate $s^{(m, i)}$. @fig-03(D) shows the relationship
between hyper-fitness and replicate fitness values for this simulation. As shown
by the posterior predictive checks examples in @fig-03(E), the hierarchical
model can correctly fit the data for each experimental replicate. Furthermore,
@fig-03(F-G) show a high degree of correlation between the ground truth and the
inferred fitness values. The ECDFs shown in @fig-03(H-I) reveal that for
$\approx 75\%$ of the non-neutral barcodes, the ground truth hyper-fitness
values fall within one standard deviation from the mean value in the posterior
distributions.

![**Hierarchical model on experimental replicates**. (A) Bayesian network
schematizing the statistical dependence between observations (filled circles)
and latent variables (empty circles). The replicates are connected via the
hyper-parameters $\underline{\theta}^M$. See `point to SI` for detailed
definitions of each variable. (B-C) Simulated replicate datasets with 900
barcodes of interest and 100 neutral lineages. (D) Comparison between the
simulation ground truth hyper-fitness and each replicate ground truth fitness.
The scatter between parameters captures experimental batch effects. (E) Examples
of the posterior predictive checks for all neutral lineages (upper left panels)
and a subset of representative mutant lineages. Shaded regions from light to
dark represent the 95\%, 68\% and 5\% credible regions. (F-G) Comparison between
the ground truth hyperfitness (F) and replicate fitness (G) values from the
simulation and the inferred parameters. Gray error bars represent the 68\%
posterior credible region. (H-I) The empirical cumulative distribution function
(ECDF) for the absolute z-score value of the ground truth parameter value within
the inferred hyper-fitness posterior distribution (H) and replicate fitness
(I).](./figs/fig03){#fig-03}

As shown in @fig-04, the structure imposed by the hierarchical model, where we
explicitly account for the connection between experimental replicates can
improve the quality of the inference.  Inferred fitness values between
experimental replicate exhibit a stronger degree of correlation in the
hierarchical model (@fig-04(A)) compared to conducting inference on each
replicate independently (@fig-04(B)). Moreover, when comparing the inferred
hyper-fitness values---the objective parameter when performing multiple
experimental measurements---the hierarchical model outperforms averaging the
independent experimental replicates as shown in @fig-04(C) and (D).

![**Comparison between hierarchical model and single dataset model**. (A-B)
comparison of inferred fitness values between experimental replicates when
fitting a hierarchical model (A) or independently fitting each dataset (B). Gray
error bars represent the 68\% credible regions. (C) Comparison between the
ground truth hyper-fitness value and the inferred parameters. The blue dots show
the inferred hyper-fitness values when assuming a hierarchical model. Gray error
bars show the 68\% credible region for this inference. The yellow dots show the
average of the mean inferred fitness values for the two experimental replicates.
No error bars are shown for these as it is inappropriate to compute one with two
data points. (D) Empirical cumulative distribution function (ECDF) of the
absolute difference between the inferred mean and the ground truth
hyper-fitness.](./figs/fig04){#fig-04}