### 2023-09-11 doc/_hierarchical_genotypes_results.qmd
continue to make progress on paper

---
editor:
    render-on-save: true
bibliography: references.bib
csl: ieee.csl
---

## Accounting for multiple barcodes per genotype via hierarchical models

![**Hierarchical model for multiple barcodes per
genotype**](./figs/fig05){#fig-05}