### 2023-09-11 doc/_variational_inference_primer.qmd

---
editor:
    render-on-save: true
bibliography: references.bib
csl: ieee.csl
---

## Primer on Variational Inference

In this section, we will briefly introduce the idea behind variational
inference. Recall that any Bayesian inference problem deals with the joint
distribution between observations $\underline{x}$ and unobserved latent
variables $\underline{\theta}$. This joint distribution can be written as the
product of a distribution of the observations $\underline{x}$ conditioned on the
$\underline{\theta}$ and the marginal distribution of these latent variables,
i.e.,
$$
\pi(\underline{x}, \underline{\theta}) =
\pi(\underline{x} \mid \underline{\theta}) \pi(\underline{\theta}).
$${#eq-vi_joint_dist}
A Bayesian inference pipeline's objective is to compute the latent variables'
posterior probability given a set of observations. This computation is
equivalent to updating our prior beliefs about the set of values that the latent
variables take after taking in new data. We write this as Bayes theorem
$$
\pi(\underline{\theta} \mid \underline{x}) = 
\frac{
        \pi(\underline{x} \mid \underline{\theta})\pi(\underline{\theta})
    }{
        \pi(\underline{x})
    }.
$${#eq-vi_bayes_theorem}
The main technical challenge for working with @eq-vi_bayes_theorem comes from
the computation of the denominator, also known as the *evidence* or the
*marginalized likelihood*. The reason computing this term is challenging is
because it involves a (potentially) high-dimensional integral of the form
$$
\pi(\underline{x}) = 
\int\cdots\int d\underline{\theta}^K \pi(\underline{x}, \underline{\theta}) = 
\int\cdots\int d\underline{\theta}^K \pi(\underline{x} \mid \underline{\theta})
\pi(\underline{\theta}),
$${#eq-vi_evidence_int}
where $K$ is the dimesionality of the $\underline{\theta}$ vector. Here, the
integrals are taken over the support---the set of values valid for the
distribution---of $\pi(\underline{\theta})$. However, only a few selected
distributions have a closed analytical form; thus, in most cases
@eq-vi_evidence_int must be solved numerically.

Integration in high-dimensional spaces can be computationally extremely
challenging. For a naive numerical quadrature procedure, integrating over a grid
of values for each dimension of $\underline{\theta}$ comes with an exponential
explosion of the number of required grid point evaluations, most of which do not
contribute significantly to the integration. To gain visual intuition about this
challenge, imagine integrating the function depicted in @fig-SI01. If the
location of the high-density region (dark peak) is unknown, numerical quadrature
requires many grid points to ensure we capture this peak. However, most of the
numerical evaluations of the function on the grid points do not contribute
significantly to the integral. Therefore, our computational resources are wasted
on insignificant evaluations. This only gets worse as the number of dimensions
increases since the number of grid point evaluation scales exponentially.

![**High-dimensional numerical quadrature does not scale with dimensionality.**
Schematic depiction of the problem with naive numerical quadrature to integrate
over an unknown density. While the density is concentrated on the dark peak,
most of the evaluations over the $x_1-x_2$ grid do not contribute to the value
of the integral](./figs/figSIX_gaussian_peak.png){#fig-SI01}

Modern Markov Chain Monte Carlo algorithms, such as Hamiltonian Monte Carlo, can
efficiently perform this high-dimensional integration by utilizing gradient
information from the target density @betancourt2017. Nevertheless, these
sampling-based methods become prohibitively slow for the number of dimensions
our present inference problem presents. Thus, there is a need to find scalable
methods for the inference problem in @eq-vi_bayes_theorem.

Variational inference circumvents these technical challenges by proposing an
approximate solution to the problem. Instead of working with the posterior
distribution in its full glory $\pi(\underline{\theta} \mid \underline{x})$, let
us propose an approximate posterior distribution $q_\phi$ that belongs to a
distribution family fully parametrized by $\phi$. For example, let us say that
the distribution $q_\phi$ belongs to the family of multivariate Normal
distributions such that $\phi = (\underline{\mu},
\underline{\underline{\Sigma}})$, where $\underline{\mu}$ is the vector of means
and $\underline{\underline{\Sigma}}$ is the covariance matrix. If we replace
$\pi$ by $q_\phi$, we want $q_\phi$ to resemble the original posterior as much
as possible. Mathematically, this can be expressed as minimizing a "*distance
metric*"---the Kullback-Leibler (KL) divergence, for example---between the
distributions. Note that we use quotation marks because, formally, the KL
divergence is not a distance metric since it is not symmetric. Nevertheless, the
variational objective is set to find a distribution $q_\phi^*$ such that
$$
q_\phi^*(\underline{\theta} \mid \underline{x}) =
\min_\phi D_{KL}\left(
    q_\phi(\underline{\theta} \mid \underline{x}) \vert\vert 
    \pi(\underline{\theta} \mid \underline{x})
\right),
$${#eq-vi_objective}
where $D_{KL}$ is the KL divergence. We alert the reader that in the literature
$q_\phi$ is usually  written not conditioned on the observations, i.e.,
$q_\phi(\underline{\theta})$. But we think that writing the conditioning
reflects that the selection of the distribution $q_\phi^*$ out of the parametric
family depends on having the observations $\underline{x}$ in hand. Furthermore,
we highlight that the KL divergence is a strictly positive number, i.e.,
$$
D_{KL}\left(
    q_\phi(\underline{\theta} \mid \underline{x}) \vert\vert 
    \pi(\underline{\theta} \mid \underline{x})
\right) \geq 0,
$${#eq-vi_kl_pos}
as this property will become important later on.

At first sight, @eq-vi_objective does not improve the situation but only
introduces further technical complications. After all, the definition of the KL
divergence
$$
D_{KL}\left(
    q_\phi(\underline{\theta} \mid \underline{x}) \vert\vert 
    \pi(\underline{\theta} \mid \underline{x})
\right) \equiv 
\int \cdots \int d\underline{\theta}^K
q_\phi(\underline{\theta} \mid \underline{x})
\ln \frac{
    q_\phi(\underline{\theta} \mid \underline{x})
}{
    \pi(\underline{\theta} \mid \underline{x})
},
$${#eq-vi_kl_div}
includes the posterior distribution $\pi(\underline{\theta} \mid \underline{x})$
we are trying to get around. However, let us manipulate @eq-vi_kl_div to beat it
to a more reasonable form. First, we can use the properties of the logarithms to
write
$$
D_{KL}\left(
    q_\phi(\underline{\theta} \mid \underline{x}) \vert\vert 
    \pi(\underline{\theta} \mid \underline{x})
\right) = 
\int d\underline{\theta}^K q_\phi(\underline{\theta} \mid \underline{x})
\ln q_\phi(\underline{\theta} \mid \underline{x}) -
\int d\underline{\theta}^K q_\phi(\underline{\theta} \mid \underline{x})
\ln \pi(\underline{\theta} \mid \underline{x}),
$${#eq-vi_step01}
where, for convenience, we write a single integration sign
($d\underline{\theta}^K$ still represents a multi-dimensional differential). For
the second term in @eq-vi_step01, we can substitute the term inside the
logarithm using @eq-vi_bayes_theorem. This results in
$$
\begin{aligned}
D_{KL}\left(
    q_\phi(\underline{\theta} \mid \underline{x}) \vert\vert 
    \pi(\underline{\theta} \mid \underline{x})
\right) &= 
\int d\underline{\theta}^K q_\phi(\underline{\theta} \mid \underline{x})
\ln q_\phi(\underline{\theta} \mid \underline{x}) \\
&- \int d\underline{\theta}^K q_\phi(\underline{\theta} \mid \underline{x})
\ln \left( 
    \frac{
        \pi(\underline{x} \mid \underline{\theta})\pi(\underline{\theta})
    }{
        \pi(\underline{x})
    }
\right).
\end{aligned}
$${#eq-vi_step02}
Again, using the properties of logarithms, we can split @eq-vi_step02, obtaining
$$
\begin{aligned}
D_{KL}\left(
    q_\phi(\underline{\theta} \mid \underline{x}) \vert\vert 
    \pi(\underline{\theta} \mid \underline{x})
\right) &= 
\int d\underline{\theta}^K q_\phi(\underline{\theta} \mid \underline{x})
\ln q_\phi(\underline{\theta} \mid \underline{x}) \\
&-\int d\underline{\theta}^K q_\phi(\underline{\theta} \mid \underline{x})
\ln \pi(\underline{x} \mid \underline{\theta}) \\
&-\int d\underline{\theta}^K q_\phi(\underline{\theta} \mid \underline{x})
\ln \pi(\underline{\theta}) \\
&+\int d\underline{\theta}^K q_\phi(\underline{\theta} \mid \underline{x})
\ln \pi(\underline{x}).
\end{aligned}
$${#eq-vi_step03}
It is convenient to write @eq-vi_step03 as
$$
\begin{aligned}
D_{KL}\left(
    q_\phi(\underline{\theta} \mid \underline{x}) \vert\vert 
    \pi(\underline{\theta} \mid \underline{x})
\right) &= 
\int d\underline{\theta}^K q_\phi(\underline{\theta} \mid \underline{x})
\ln \frac{
    q_\phi(\underline{\theta} \mid \underline{x})
    }{
        \pi(\underline{\theta})
    } \\
&-\int d\underline{\theta}^K q_\phi(\underline{\theta} \mid \underline{x})
\ln \pi(\underline{x} \mid \underline{\theta}) \\
&+ \ln \pi(\underline{x}) 
\int d\underline{\theta}^K q_\phi(\underline{\theta} \mid \underline{x}),
\end{aligned}
$${#eq-vi_step04}
where for the last term, we can take $\ln \pi(\underline{x})$ out of the
integral since it does not depend on $\underline{\theta}$. Lastly, we utilize
two properties:

1. The proposed approximate distribution must be normalized, i.e.,
$$
\int d\underline{\theta}^K q_\phi(\underline{\theta} \mid \underline{x}) = 1.
$${#eq-vi_q_norm}
2. The law of the unconscious statistician (LOTUS) establishes that for any
probability density function, it must be true that
$$
\int d\underline{\theta}^K q_\phi(\underline{\theta} \mid \underline{x})
f(\underline{\theta}) = \left\langle 
    f(\underline{\theta}) 
\right\rangle_{q_\phi},
$${#eq-vi_lotus}
where $\left\langle\cdot\right\rangle_{q_\phi}$ is the expected value over the
$q_\phi$ distribution.

Using these two properties, the positivity constraint on the KL divergence in
@eq-vi_kl_pos, and the definition of the KL divergence in @eq-vi_kl_div we can
rewrite @eq-vi_step04 as
$$
D_{KL}\left( 
    q_\phi(\underline{\theta} \mid \underline{x}) \vert \vert
    \pi(\underline{\theta}) 
\right) -
\left\langle
    \ln \pi(\underline{x} \mid \underline{\theta})
\right\rangle_{q_\phi}
\geq - \ln \pi(\underline{x}).
$${#eq-vi_step05}
Multiplying by a minus one, we have the functional form of the so-called
evidence lower bound (ELBO) @kingma2014,
$$
\underbrace{
    \ln \pi(\underline{x})
}_{\text{log evidence}} \geq
\underbrace{
    \left\langle
        \ln \pi(\underline{x} \mid \underline{\theta})
    \right\rangle_{q_\phi} -
    D_{KL}\left( 
        q_\phi(\underline{\theta} \mid \underline{x}) \vert \vert
        \pi(\underline{\theta}) 
    \right)
}_{\text{ELBO}}.
$${#eq-vi_elbo}

Let us recapitulate where we are. We started by presenting the challenge of
working with Bayes' theorem, as it requires a high-dimensional integral of the
form in @eq-vi_evidence_int. As an alternative, variational inference posits to
approximate the posterior distribution $\pi(\underline{\theta} \mid
\underline{x})$ with a parametric distribution $q_\phi(\underline{\theta} \mid
\underline{x})$. By minimizing the KL divergence between these distributions, we
arrive at the result in @eq-vi_elbo, where the left-hand side---the log
marginalized likelihood or log evidence---we cannot compute for
technical/computational reasons. However, the right-hand side is composed of
things we can easily evaluate. We can easily evaluate the log-likelihood $\ln
\pi(\underline{x} \mid \underline{\theta})$ and the KL divergence between our
proposed approximate distribution $q_\phi(\underline{\theta} \mid
\underline{x})$ and the prior distribution $\pi(\underline{\theta})$. Moreover,
we can compute the gradients of these functions with respect to the parameters
of our proposed distribution. This last point implies that we can change the
parameters of the proposed distribution to maximize the ELBO. And, although we
cannot compute the left-hand side of @eq-vi_elbo, we know that however large we
make the ELBO, it will always be smaller than (or equal) the log-marginal
likelihood. Therefore, the larger we can make the ELBO by modifying the
parameters $\phi$, the closer it gets to the log-marginal likelihood, and, as a
consequence, the better our proposed distribution $q_\phi(\underline{\theta}
\mid \underline{x})$ gets to the true posterior distribution
$\pi(\underline{\theta} \mid \underline{x})$.

In this sense, variational inference turns the intractable numerical integration
problem to an optimization routine, for which there are several algorithms 
available.

### ADVI algorithm

To maximize the right-hand side of @eq-vi_elbo, the Automatic Differentiation
Variational Inference (ADVI) algorithm developed in [@kucukelbir2016] takes
advantage of advances in probabilistic programming languages to generate a
robust method to perform this optimization. Without going into the details of
the algorithm implementation, for our purposes, it suffices to say that we define
our joint distribution $\pi(\underline{\theta}, \underline{x})$ as the product
defined in @eq-vi_joint_dist. ADVI then proposes an approximate variational
distribution $q_\phi$ that can either be a multivariate Normal distribution with
a diagonal covariance matrix, i.e.,
$$
\phi = (\underline{\mu}, \underline{\underline{I}}\;\underline{\sigma}^2),
$${#eq-vi_meanfield}
where $\underline{\underline{I}}$ is the identity matrix, and
$\underline{\sigma}$ is the vector of standard deviations or a full-rank 
multivariate Normal distribution
$$
\phi = (\underline{\mu}, \underline{\underline{\Sigma}}).
$${#eq-vi_full}

Then, the parameters are initialized in some value $\phi_o$. These parameters
are iteratively updated by computing the gradient of the ELBO (right-hand side
of @eq-vi_elbo), hereafter defined as $\mathcal{L}$, with respect to the
parameters, 
$$
\nabla_\phi \mathcal{L} = \nabla_{\underline{\mu}} \mathcal{L} + 
\nabla_{\underline{\sigma}}\mathcal{L},
$${#eq-vi_elbo_gradient}
and then computing
$$
\phi_{t+1} = \phi_{t} + \eta \nabla_\phi \mathcal{L},
$$
where $\eta$ defines the step size.

This short explanation behind the ADVI algorithm is intended only to gain
intuition on how the optimal variational distribution $q_\phi$ be computed.
There are many nuances in the implementation of the ADVI algorithm. We invite
the reader to look at the original reference for further details.