### 2023-09-13 doc/_experiment.qmd

---
editor:
    render-on-save: true
bibliography: references.bib
csl: ieee.csl
---

## Experimental setup{#sec-experiment}

The present work is designed to analyze time-series data of relative abundance
of multiple microbial lineages uniquely identified by a DNA barcode
[@kinsler2020; @ascensao2023]. In these competition assays, an ensemble of
genotypes is pooled together with an unlabeled reference strain that, initially,
represents the vast majority ($\geq 90\%$) of the cells in the culture (see
schematic in @fig-01(A)). Furthermore, a fraction of labeled genotypes
equivalent to the unlabeled reference strain---hereafter defined as *neutral*
lineages---are spiked in at a relatively high abundance ($\approx 3-5\%$). The
rest of the culture is left for the ensemble of genotypes of interest.

To determine the relative fitness of the ensemble of genotypes, a series of
growth-dilution cycles are performed on either a single or multiple
environments. In other words, the cultures are grown for some time to then
dilute the culture back into fresh media and repeat for roughly 4-7 cycles,
depending on the initial abundances of the mutants, and their relative growth
rates. The DNA barcodes are sequenced at the end of each growth cycle to
quantify the relative abundance of each of the barcodes. We point the reader to
@kinsler2020 for specific details on these assays for *S. cerevisiae* and to
@ascensao2023 for equivalent assays for *E. coli*. @fig-01(B) presents a typical
barcode trajectory where the darker barcode trajectories represent the so-called
*neutral lineages* that are genetically equivalent to the untagged ancestor
strain that initially dominates the culture. These neutral lineages simplify the
inference problem since the fitness of all relevant barcodes is quantified with
respect to these barcodes.

![**Typical competitive fitness experiment**. (A) Schematic of the typical
experimental design to determine the competitive fitness of an ensemble of
barcoded genotypes. Genotypes are pooled together and grown over multiple
growth-dilution cycles. At the end of each cycle, a sample is processed to
generate a library for amplicon sequencing. (B) Typical barcode trajectory
dataset. From each time point, the relative frequency of each barcode is
determined from the total number of reads. Different shades of blue represent
different relative fitness. Darker gray lines define the typical trajectory of
neutral lineages. (C) Bayesian network defining the statistical dependence
between observations (filled circles) and latent variables (empty circles). The
dotted circle highlights the set of variables we are interested in. One-way
arrows depict the statistical dependence between variables. For example, the
value of $\underline{\underline{F}}^N$ only depends on the value of
$\underline{\bar{s}}_T$, while $\underline{\underline{F}}^M$ depends on
$\underline{\bar{s}}_T$ and $\underline{s}^M$. See @sec-bayesian_inference for
specific definitions of each variable.](./figs/fig01){#fig-01}