### 2023-09-13 doc/_intro.qmd

---
editor:
    render-on-save: true
bibliography: references.bib
csl: ieee.csl
---

The advent of DNA barcoding---the ability to uniquely identify cell lineages
with DNA sequences integrated at a specific locus---and high-throughput
sequencing has opened new venues for understanding microbial evolutionary
dynamics with an unprecedented level of temporal resolution [@levy2015;
@nguyenba2019a; @ascensao2023]. These experimental efforts rely on our ability
to reliably infer the relative fitness of an ensemble of diverse genotypes.
Moreover, inferring these fitness values over an ensemble of environmental
conditions can help us determine the phenotypic diversity of a rapid adaptation
process [@kinsler2020].

As with any other sequencing-based quantification, tracking lineages via DNA
barcode sequencing is inexorably accompanied by noise sources coming from
experimental manipulation of the microbial cultures, DNA extraction and
sequencing library preparation that involves multiple rounds of PCR
amplification, and the sequencing process itself. Thus, accounting for the
uncertainty when inferring the relevant parameters from the data is a key step
to draw reliable conclusions. Bayesian statistics presents a paradigm by which
one can account for all known sources of uncertainty in a principled way
[@eddy2004a]. This, combined with the development of modern Markov Chain Monte
Carlo sampling algorithms [@betancourt2017] and approximate variational
approaches [@kucukelbir2016] have boosted a resurgence of Bayesian methods in
different fields [@efron2013a].

We present a Bayesian inference pipeline to properly quantify the uncertainty
about the parametric information we can extract from high-throughput competitive
fitness assays. In these assays, the fitness of an ensemble of genotypes is
determined relative to a reference genotype [@kinsler2020; @ascensao2023].
@fig-01(A) shows a schematic of the experimental procedure in which an initial
pool of barcoded strains are mixed with a reference strain and inoculated into
fresh media. After some time---usually, enough time for the culture to
saturate---an aliquot is transferred to fresh media, while the remaining culture
is used for DNA sequencing of the lineage barcodes. The time-series information
of the relative abundance of each lineage, i.e., the barcode frequency depicted
in @fig-01(B), is used to infer the relative fitness---the growth advantage on a
per-cycle basis---for each lineage with respect to the reference strain. The
proposed statistical model accounts for multiple sources of uncertainty when
inferring the lineages' relative fitness values (see @sec-experiment for details
on sources of uncertainty accounted by the model). Furthermore, small changes to
the core statistical model allow us to account for relevant experimental
variations of these competition assays. More specifically, as described in
@sec-replicates, our statistical model can account for batch-to-batch
differences when jointly analyzing multiple experimental replicates using a
Bayesian hierarchical model. Moreover, a variant of these hierarchical models,
presented in @sec-genotypes, can account for variability within multiple
barcodes mapping to equivalent genotypes within the same experiment. Finally, in
@sec-multienv, we present a variation of the statistical model to infer fitness
on growth-dilution cycles in multiple environments.

For all the model variations presented in this paper, we benchmark the ability
of our pipeline to infer relative fitness parameters against synthetic data with
known ground truth. We also re-analyze published data from @kinsler2020 to
assess the method's performance on experimental measurements. The present method
is accompanied by a `Julia` package to readily implement the inference pipeline
with state-of-the-art scientific computing software. [`here mention something
about the Python wrapper if appropriate`].