### 2023-09-13 doc/_multienv_results.qmd

---
editor:
    render-on-save: true
bibliography: references.bib
csl: ieee.csl
---

## Fitness inference on multiple environments{#sec-multienv}

![**Multi-environment fitness inference.**](./figs/fig06){#fig-06}