### 2023-09-13 doc/_single_results.qmd

---
editor:
    render-on-save: true
bibliography: references.bib
csl: ieee.csl
---

## Inference on a single dataset

To assess the inference pipeline performance, we applied it to a simulated
dataset with known ground truth relative fitness values (See `[ref to SI]` for
details on simulation). @fig-02(A) shows the structure of the synthetic dataset.
The majority of barcodes of interest (faint color lines) are adaptive compared
to the neutral barcodes ($s^{(m)} > 0$). This experimental condition is required
to reliably measure the relative fitness since all barcodes of interest occupy a
small fraction of the population at the beginning of the experiment `[ref SI
section on detrimental mutants]`. Although the barcode frequency trajectories
look relatively smooth, our fitness model requires the computation of the log
frequency ratio between adjacent time points as derived in @eq-logfreq.
@fig-02(B) shows such data transformation where we can better appreciate the
observational noise input into our statistical model. This noise is evident for
the darker lines representing the neutral barcodes since all of these lineages
are assumed to be identically distributed under this log frequency ratio
transformation.

To visualize the performance of our inference pipeline in fitting our fitness
model to the observed data, we compute the so-called posterior predictive checks
(PPC). In short, the PPC consists of repeatedly generating synthetic datasets in
agreement with the results from the inference results. In other words, we use
the resulting parameter values from the ADVI inference to generate possible
datasets in agreement with the inferred values. @fig-02(C) shows these results
for all neutral lineages (upper left corner plot) and a few representative
non-neutral barcodes. The different color shades represent the 95\%, 68\%, and
5\% credible regions, i.e., the regions where we expect to find the data with
the corresponding probability---or in terms of our parameter, the $X\%$ credible
region is the interval where we expect the true parameter to lie with $X\%$
probability.

The main advantage of our method is this natural interpretability of these
credible regions where an $X\%$ credible region indeed captures the region of
parameter space where we expect with $X\%$ probability the actual value of the
parameter lies given our statistical model, our prior information, and the
observed experimental data. A common mistake in the literature is interpreting
frequentist confidence intervals as Bayesian credible regions when they are not
equivalent [@morey2016]. Frequentist confidence intervals and Bayesian credible
regions are based on fundamentally different philosophical approaches to
statistics. Frequentist confidence intervals represent the range of values that
would contain the true population parameter with a certain probability if the
experiment was repeated many times. The confidence level does not represent the
probability that the interval contains the true value. According to a specific
model and prior information, Bayesian credible regions represent the range of
values that contain the parameter with a certain posterior probability. The
credible region directly represents the probability that the region contains the
true value. So, frequentist confidence intervals cannot be interpreted as
Bayesian credible regions because they have fundamentally different meanings.
Treating an $X\%$ confidence interval like an $X\%$  credible region is
fallacious since confidence levels do not represent probabilistic coverage of
the true value like credible regions. The intervals are generated through
entirely different procedures.

To capture the global performance of the model, @fig-02(D) compares the known
ground truth with the inferred relative fitness value for all barcodes of
interest. There is an excellent degree of correspondence between these values,
with the error bars representing the 68\% credible region for the parameter
value crossing the identity line for most barcodes. This latter point is made
clear with @fig-02(E) where $\approx 90\%$ of ground truth fitness values fall
within one standard deviation of the mean in the inferred posterior
distributions.

![**Single dataset inference**. (A) Frequency trajectories that represent the
raw data going into the inference. (B) Log frequency ratio between two adjacent
time points used by the inference pipeline. Darker lines represent the neutral
barcodes. These transformed data are much more noisy than the
seemingly smooth frequency trajectories. (B) Examples of the posterior
predictive checks for all neutral lineages (upper left panel) and a subset of
representative mutant lineages. Error bars represent the 68\% credible region on
the mutant relative fitness marginal distribution. (C) Comparison between the
ground truth fitness value from the logistic-growth simulation and the inferred
fitness value. Gray error bars represent the 68\% posterior credible region. (D)
The empirical cumulative distribution function (ECDF) for the absolute z-score
value of the ground truth parameter value within the inferred fitness
posterior distribution.](./figs/fig02){#fig-02}