### 2023-09-13 doc/_multienv_results.qmd
first draft of multi-env section

---
editor:
    render-on-save: true
bibliography: references.bib
csl: ieee.csl
---

## Fitness inference on multiple environments{#sec-multienv}

One of the benefits of Bayesian methods is the intrinsic error propagation
embedded in the mathematical framework. For example, the uncertainty on the
value of the population mean fitness values $\underline{\bar{s}}_T$ is
propagated to the relative fitness of a particular strain $s^{(b)}$ since we
defined a joint posterior distribution over all parameters as fully expressed in
@eq-bayes_full. This natural error propagation can help us with the experimental
design schematized in @fig-06(A). Here, rather than performing growth-dilution
cycles in the same environment, the cells are diluted into a different 
environment. Thus, the uncertainty on the fitness estimate for the previous
environment must be propagated to that of the next one.

To validate the extension of our statistical model to this scenario, @fig-06(B)
shows the trajectory of the log frequency ratios between adjacent time points.
The different colored regions correspond to the different environments. For this
simulation, the growth rate of Environment 2 was set to be, on average, half of
the average growth rate in Environment 1. Equivalently, the growth rate in
Environment 3 was set to be, on average, twice the average growth rate in
Environment 1. @fig-06(C-E) show the correspondence between the simulation
ground truth and the inferred fitness values, where the error bars represent the
68\% credible region. @fig-06(F) summarizes the performance of our inference
pipeline by showing the empirical cumulative distribution functions for the
absolute value of the ground truth fitness value z-score within the posterior
distribution. This plot shows that, overall, $\approx 75\%$ of inferred mean
values fall within one standard deviation of the ground truth. For completeness,
@fig-06(G) shows the posterior predictive checks for a few example barcodes.

![**Multi-environment fitness inference.** (A) Schematic of the simulated
experimental design where growth-dilution cycles were performed in different
environments in a cycle. (B) log frequency rations between adjacent time points.
Darker lines represent the neutral barcodes. The colors in the background demark
the corresponding environment, matching colors in (A). (C-E) Comparison between
the ground truth fitness value from the logistic-growth simulation and the
inferred fitness value for each environment. Gray error bars represent the 68\%
posterior credible region. (F) The empirical cumulative distribution function
(ECDF) for the absolute z-score value of the ground truth parameter value within
the inferred fitness posterior distribution for all fitness values (black line)
and each environment individually (color lines). (G) Examples of the posterior
predictive checks for all neutral lineages (upper left panel) and a subset of
representative mutant lineages. Shaded regions represent the 95\%, 68\%, and 5\%
credible regions for the data. The reported errors above the plot represent the
68\% credible region on the mutant relative fitness marginal distribution.
Background colors match those of (A).](./figs/fig06){#fig-06}