### 2023-09-15 doc/_hierarchical_genotypes_results.qmd

---
editor:
    render-on-save: true
bibliography: references.bib
csl: ieee.csl
---

## Accounting for multiple barcodes per genotype via hierarchical models{#sec-genotypes}

Hierarchical models can also capture another experimental design in which
multiple barcodes map to the same or an equivalent genotype. As we will show,
this many-to-one mapping can improve the inference compared to the extreme cases
of inferring the fitness of each barcode independently or pooling the data of
all barcodes mapping to a single genotype. As schematized in @fig-05(A), a small
modification allows us to map the structure of our original model to that of a
hierarchical model with a fitness hyperparameter vector $\underline{\theta}^G$,
where $G$ is the number of genotypes in the dataset.

@fig-05(B) shows a single experimental replicate in which 90 genotypes were
assigned a random number of barcodes (a multinomial distribution with a mean of
ten barcodes per genotype) for a total of 900 non-neutral barcodes. To assess
the performance of the hierarchical model proposed in @fig-05(A), we performed
inference using this hierarchical model, as well as the two extreme cases of
ignoring the connection between the barcodes belonging to the same genotype
---equivalent to performing inference using the model presented in @fig-02(A)
over the barcodes---or pooling the data of all barcodes belonging to the same
genotype into a single count---equivalent to performing inference using the
model presented in @fig-02(A) over the pooled barcodes. @fig-05(C-D) shows the
comparison between the simulation ground truth and the inferred values for these
three cases. Not only do the hierarchical model results show higher degrees of
correlation with the ground truth, but the error bars (representing the 64\%
credible regions) are smaller, meaning that the uncertainty in the estimate
decreases when using the hierarchical model. The improvement in the prediction
can be seen in @fig-05(F) where the empirical cumulative distribution function
of the absolute difference between the mean inferred value and the simulation
ground truth is shown for all three inference models. The hierarchical model's
curve ascends more rapidly, showing that, in general, the inferred values are
closer to the ground truth. For completeness, @fig-05(G) shows some examples of
how the hierarchical model can capture the raw log-frequency count observations.

![**Hierarchical model for multiple barcodes per genotype.** (A) Schematic
depiction of the hierarchical structure for multiple barcodes mapping to a
single genotype. A set of barcodes mapping to an equivalent genotype map to
"local" fitness values $s^{(b)}$ that are connected via a hyper-fitness
parameter for the genotype $\theta^{(g)}$. (B) Simulated dataset with 100
neutral lineages and 900 barcodes of interest distributed among 90 genotypes.
(C-E) Comparison between the inferred and ground truth fitness values for a
hierarchical model (C), a model where each barcode is inferred independently
(D), and a model where barcodes mapping to the same genotype are pooled
together (E). Gray error bars represent the 68\% credible regions. (F) Empirical
cumulative distribution function (ECDF) of the absolute difference between the
inferred mean and the ground truth fitness values for all three models. (G)
Examples of the posterior predictive checks for all neutral lineages (upper left
panels) and a subset of representative mutant lineages. Shaded regions from
light to dark represent the 95\%, 68\%, and 5\% credible
regions.](./figs/fig05){#fig-05}