### 2023-09-19 doc/_logistic_sim.qmd
add logistic growth section

---
editor:
    render-on-save: true
bibliography: references.bib
csl: ieee.csl
---

## Logistic growth simulation

In this section, we explain the simulations used to assess the validity of our
inference pipeline. Let us begin by assuming that, since the strains are grown
for two full days in the experiment, having left behind the exponential phase
for almost an entire day, a simple exponential growth of the form
$$
\frac{dn_i}{dt} = \lambda_i n_i,
$${#eq-logistic_expo}
where $n_i$ is the number of cells of strain $i$, and $\lambda_i$ is the
corresponding growth rate is not enough. Instead, we will assume that the cells
follow the logistic growth equation of the form
$$
\frac{dn_i}{dt} = \lambda_i n_i 
\left( 1 - \frac{\sum_{j=1}^{N}n_i}{\kappa}\right),
$${#eq-logistic_fn}
where $\kappa$ is the carrying capacity, and $N$ is the total number of strains
in the culture.

The inference method is based on the model that assumes that the time passed
between dilutions $\tau \approx 8$ generations, the change in frequency for a
mutant barcode can be approximated from cycle $t$ to the next cycle $t + 1$ as
$$
f_{t+1}^{(m)} = f_t^{(m)} \mathrm{e}^{(s^{(m)} - \bar{s}_t)\tau},
$${#eq-logistic_fitness_fn}
where $s^{(m)}$ is the relative fitness for strain $i$ compared to the ancestral
strain and $\bar{s}_t$ is the mean fitness of the population at cycle $t$. We
implemented a numerical experiment following the logistic growth model described
in @eq-logistic_fn to test this assumption. @fig-SIX_logistic_growth_01 shows an
example of the deterministic trajectories for 50 labeled neutral lineages and
1000 lineages of interest. The upper red curve that dominates the culture 
represents the unlabeled ancestral strain included in the experimental design
described in @sec-experiment.

![**Logistic growth simulation over single growth cycle**. The dashed line
represents the neutral lineages, with the upper curve being the unlabeled
neutral strain. Color curves represent the genotypes of interest colored by
growth rate relative to the neutral
lineage.](./figs/figSIX_logistic_growth_01){#fig-SIX_logistic_growth_01}

To simulate multiple growth-dilution cycles, we take the population composition
at the final time point and use it to initialize a new logistic growth
simulation. @fig-SIX_logistic_growth_02 shows the resulting number of cells at
the final time point of a cycle over multiple growth-dilution cycles for the
genotypes in @@fig-SIX_logistic_growth_01. We can see that the adaptive lineages
(blue curves) increase in abundance, while detrimental lineages (red curves)
decrease.

![**Growth-dilution cycles for logistic growth simulation**. Each point
represents the final number of cells after a growth cycle for each lineage.
Colors are the same as in
@fig-SIX_logistic_growth_01.](./figs/figSIX_logistic_growth_02){#fig-SIX_logistic_growth_02}

In @sec-fitness_model we derive the functional form to infer the relative 
fitness of each lineage as
$$
\frac{1}{\tau}\ln \frac{f_{t+1}^{(b)}}{f_{t}^{(b)}} = (s^{(b)} - \bar{s}_t).
$${#eq-logistic_logfreq}
@fig-SIX_logistic_growth_03 shows the corresponding log frequency ratio curves
for the logistic growth simulation. The displacement of these curves with 
respect to the neutral lineages determines the ground truth relative fitness
value for these simulations.

![**Log frequency ratio for logistic growth simulations**. The relative distance
of the color curves from the black dashed line determines the relative fitness
of each linage.](./figs/figSIX_logistic_growth_03){#fig-SIX_logistic_growth_03}

To simulate the experimental noise, we add two types of noise:

1. Poisson noise between dilutions. For this, we take the final point of the
logistic growth simulation and sample a random Poisson number based on this
final point to set the initial condition for the next cycle.
2. Gaussian noise when performing the measurements. When translating the 
underlying population composition to the number of reads, we can add a custom 
amount of Gaussian noise.

@fig-SIX_logistic_growth_04 shows the frequency trajectories (left panels) and
log frequency ratios (right panels) for a noiseless simulation (upper panels) 
and a simulation with added noise (lower panels). The noiseless simulation is
used to determine the relative fitness for each of the lineages, which serves
as the ground truth to be compared with the resulting inference.

![**Logistic growth-dilution simulations with and without
noise**](./figs/figSIX_logistic_growth_04){#fig-SIX_logistic_growth_04}