### 2023-09-20 doc/_bayes_hierarchical.qmd
add first draft for hierarchical model SI

---
editor:
    render-on-save: true
bibliography: references.bib
csl: ieee.csl
---

### Hierarchical models for multiple experimental replicates

As detailed in @sec-replicates of the main text, we define a Bayesian
hierarchical model to analyze data from multiple experimental replicates. The
implementation requires only slightly modifying the base model detailed in the
previous sections. The hierarchical model defines a hyper-fitness parameter
$\theta^{(m)}$ for every non-neutral barcode. We can thus collect all of the $M$
hyperparameters in an array of the form
$$
\underline{\theta}^M = (\theta^{(1)}, \ldots, \theta^{(M)})^\dagger.
$${#eq-hier_hyper_vector}
Our data now consists of a series of matrices $\underline{\underline{R}}_{[j]}$,
where the subindex $[j]$ refers to the $j$-th experimental replicate. These
matrices need not have the same number of rows, as the time points measured for
each replicate can vary. The statistical model we must define is then of the
form
$$
\begin{aligned}
\pi(
    \underline{\theta}^M, 
    \{\underline{s}^M_{[j]}\},
    \{\underline{\bar{s}}_{T[j]}\},
    \{\underline{\underline{F}}_{[j]}\} \mid
    \{\underline{\underline{R}}_{[j]}\}
) \propto \;
&\pi(
    \{\underline{\underline{R}}_{[j]}\} \mid
    \underline{\theta}^M, 
    \{\underline{s}^M_{[j]}\},
    \{\underline{\bar{s}}_{T[j]}\},
    \{\underline{\underline{F}}_{[j]}\} 
) \times \\\
&\pi(
    \underline{\theta}^M, 
    \{\underline{s}^M_{[j]}\},
    \{\underline{\bar{s}}_{T[j]}\},
    \{\underline{\underline{F}}_{[j]}\} 
)
\end{aligned}
$${#eq-hier_bayes_full}
where the parameters within curly braces with subindex $[j]$ indicate one set of
parameters per experimental replicate. For example,
$$
\{\underline{s}^M_{[j]}\} = \{
    \underline{s}^M_{[1]}, \underline{s}^M_{[2]}, \ldots \underline{s}^M_{[E]}
\},
$${#eq-hier_array}
where $E$ is the number of experimental replicates.

Given the dependencies between the variables, we can factorize
@eq-hier_bayes_full to be of the form
$$
\begin{aligned}
\pi(
    \underline{\theta}^M, 
    \{\underline{s}^M_{[j]}\},
    \{\underline{\bar{s}}_{T[j]}\},
    \{\underline{\underline{F}}_{[j]}\} \mid
    \{\underline{\underline{R}}_{[j]}\}
) =
&\pi(
    \underline{\theta}^M, 
    \{\underline{s}^M_{[j]}\} \mid
    \{\underline{\bar{s}}_{T[j]}\},
    \{\underline{\underline{F}}_{[j]}\} 
) \times \\
&\pi(
    \{\underline{\bar{s}}_{T[j]}\} \mid
    \{\underline{\underline{F}}_{[j]}\}
) \times \\
&\pi(
    \{\underline{\underline{F}}_{[j]}\} \mid
    \{\underline{\underline{R}}_{[j]}\}
)
\end{aligned}
$${#eq-hier_factorized}
Furthermore, the hierarchical structure only connects the replicates via the
relative fitness parameters. This means that the population mean fitness values
and the frequencies can be independently inferred for each dataset. This allows
us to rewrite the right-hand side of @eq-hier_factorized as
$$
\begin{aligned}
\pi(
    \underline{\theta}^M, 
    \{\underline{s}^M_{[j]}\},
    \{\underline{\bar{s}}_{T[j]}\},
    \{\underline{\underline{F}}_{[j]}\} \mid
    \{\underline{\underline{R}}_{[j]}\}
) =
&\pi(
    \underline{\theta}^M, 
    \{\underline{s}^M_{[j]}\} \mid
    \{\underline{\bar{s}}_{T[j]}\},
    \{\underline{\underline{F}}_{[j]}\} 
) \times \\
&\prod_{j=1}^E \left[
    \pi(
        \underline{\bar{s}}_{T[j]} \mid
        \underline{\underline{F}}_{[j]}
    )
    \pi(
        \underline{\underline{F}}_{[j]} \mid
        \underline{\underline{R}}_{[j]}
    )
\right].
\end{aligned}
$${#eq-hier_bayes_factorized_replicate}
The terms inside the square brackets in @eq-hier_bayes_factorized_replicate take
the same functional form as those derived in @sec-bayes_meanfit and
@sec-bayes_freq. Therefore, to implement the desired hierarchical model, we only
need to focus on the first term on the right-hand side of
@eq-hier_bayes_factorized_replicate. A way to think about the structure of the
hierarchical model is as follows: imagine each genotype as a "*true*" relative
fitness value. However, every time we perform an experiment, small variations in
the biotic and abiotic conditions---also known as batch effects---might result
in small deviations from this value. We model this by defining a distribution
for the hyper-fitness parameter---the ground truth we are interested in---and
having each experimental replicate sample from this hyper-parameter distribution
to determine the "*local*" fitness value. The wider the hyper-parameter
distribution is the more variability between experimental replicates.

Writing Bayes' theorem for the first term in @eq-hier_bayes_factorized_replicate
results in
$$
\pi(
    \underline{\theta}^M, 
    \{\underline{s}^M_{[j]}\} \mid
    \{\underline{\underline{F}}_{[j]}\},
    \{\underline{\bar{s}}_{T[j]}\}
) \propto
\pi(
    \{\underline{\underline{F}}_{[j]}\} \mid
    \underline{\theta}^M, 
    \{\underline{s}^M_{[j]}\},
    \{\underline{\bar{s}}_{T[j]}\}
)
\pi(
    \underline{\theta}^M, 
    \{\underline{s}^M_{[j]}\} \mid
    \{\underline{\bar{s}}_{T[j]}\}
),
$${#eq-hier_bayes}
where we leave the conditioning on the population mean fitness as we did in
@seq-bayes_mutfit. This expression can be simplified in two ways. First, the
frequency values for each experimental replicate depend directly on the local
fitness values and the corresponding population mean fitness, as the
relationship between experimental replicates only occurs through the relative
fitness values. Therefore, we can write
$$
\pi(
    \underline{\theta}^M, 
    \{\underline{s}^M_{[j]}\} \mid
    \{\underline{\underline{F}}_{[j]}\},
    \{\underline{\bar{s}}_{T[j]}\}
) \propto
\prod_{j=1}^E\left[
    \pi(
        \underline{\underline{F}}_{[j]} \mid
        \underline{s}^M_{[j]},
        \underline{\bar{s}}_{T[j]}
    )
\right]
\pi(
    \underline{\theta}^M, 
    \{\underline{s}^M_{[j]}\} \mid
    \{\underline{\bar{s}}_{T[j]}\}
).
$${#eq-hier_bayes_exp_indep}
Second, the relationship between the hyper-fitness and the local fitness values
allows us to write their joint distribution as a conditional distribution where
local fitness values depend on the global hyper-fitness value, obtaining
$$
\pi(
    \underline{\theta}^M, 
    \{\underline{s}^M_{[j]}\} \mid
    \{\underline{\underline{F}}_{[j]}\},
    \{\underline{\bar{s}}_{T[j]}\}
) \propto
\prod_{j=1}^E\left[
    \pi(
        \underline{\underline{F}}_{[j]} \mid
        \underline{s}^M_{[j]},
        \underline{\bar{s}}_{T[j]}
    )
    \pi(
        \underline{s}^M_{[j]} \mid
        \underline{\theta}^M
    )
\right] 
\pi(
    \underline{\theta}^M
).
$${#eq-hier_bayes_exp_indep}
Notice we removed the conditioning on the population mean fitness as our prior
expectations of what the global hyper-fitness or local fitness value might be do
not depend on these nuisance parameters.

The first term on the right-hand side of @eq-hier_bayes_exp_indep takes the same
functional form as the one derived in @seq-bayes_mutfit. Therefore, all we are
left with is to determine the functional forms for the hyper–prior
$\pi(\underline{\theta}^M)$, and the conditional probability
$\pi(\underline{s}^M_{[j]} \mid \underline{\theta}^M)$. In analogy to the 
assumptions used for the fitness values in @seq-bayes_mutfit, we define the 
value of each hyper-fitness as independent. This means that we have
$$
\pi(\underline{\theta}^M) = \prod_{m=1}^M \pi(\theta^{(m)}).
$${#eq-hier_hyper_prior}
Furthermore, we assume this prior is of the form
$$
\theta^{(m)} \sim \mathcal{N}(\mu_{\theta^{(m)}}, \sigma_{\theta^{(m)}}),
$${#eq-hier_hyper_prior_02}
where $\mu_{\theta^{(m)}}$ and $\sigma_{\theta^{(m)}}$ are user-defined 
parameters encoding the prior expectations on the fitness values.

For the conditional distribution $\pi(\underline{s}^M_{[j]} \mid
\underline{\theta}^M)$, we use the so-called non-centered parametrization that
avoids some of the intrinsic degeneracies associated with hierarchical models
[@betancourt2013]. We invite the reader to check [this excellent
blog](https://betanalpha.github.io/assets/case_studies/hierarchical_modeling.html)
explaining the difficulties of working with hierarchical models. This 
non-centered parameterization implies that we introduce two nuisance parameters
such that the local fitness $s_{[j]}^{(m)}$ is computed as
$$
s_{[j]}^{(m)} = \theta^{(m)} + 
(\tau_{[j]}^{(m)} \times \xi_{[j]}^{(m)}),
$${#eq-hier_noncentered}
where $\theta^{(m)}$ is the corresponding genotype hyper-fitness value,
$\xi_{[j]}^{(m)}$ is a standard normal random variable, i.e.,
$$
\xi_{[j]}^{(m)} \sim \mathcal{N}(0, 1),
$${#eq-hier_stdnormal}
that allows deviations from the hyper-fitness value to be either positive or
negative, and $\tau_{[j]}^{(m)}$ is a strictly positive random variable that
characterizes the deviation of the local fitness value from the global
hyper-fitness. We assume
$$
\tau_{[j]}^{(m)} \sim \log\mathcal{N}(
    \mu_{\tau_{[j]}^{(m)}}, \sigma_{\tau_{[j]}^{(m)}}
)
$${#eq-hier_tau}
where $\mu_{\tau_{[j]}^{(m)}}$ and $\sigma_{\tau_{[j]}^{(m)}}$ are user-defined
parameters capturing the expected magnitude of the batch effects.