### 2023-09-26 doc/_fitness_model.qmd
update from time to cycle number

---
editor:
    render-on-save: true
bibliography: references.bib
csl: ieee.csl
---

## Preliminaries on mathematical notation
Before jumping directly into the Bayesian inference pipeline, let us establish
the mathematical notation used throughout this paper. We define
(column) vectors as underlined lowercase symbols such as
$$
\underline{x} = \begin{bmatrix}
    x_1\\
    x_2\\
    \vdots\\
    x_N
\end{bmatrix}.
$${#eq-vec_def}
In the same way, we define matrices as double-underline uppercase symbols such
as
$$
\underline{\underline{A}} =
\begin{bmatrix}
    A_{11} & A_{12} & \cdots & A_{1N}\\
    A_{21} & A_{22} & \cdots & A_{2N}\\
    \vdots & \vdots & \ddots & \vdots\\
    A_{M1} & A_{M2} & \cdots & A_{MN}\\
\end{bmatrix}.
$${#eq-mat_def}

## Fitness model {#sec-fitness_model}

Empirically, the barcode relative frequency trajectories follow an exponential
function of the form [@levy2015; @kinsler2020; @ascensao2023]
$$
f_{t+1}^{(b)} = f_{t}^{(b)} \mathrm{e}^{(s^{(b)} - \bar{s}_t)\tau},
$${#eq-fitness}
where $f_{t}^{(b)}$ is the frequency of barcode $b$ at the end of cycle number
$t$, $s^{(b)}$ is the relative fitness with respect to the reference
strain---the quantity we want to infer from the data---$\bar{s}_t$ is the mean
fitness of the culture at the end of cycle number $t$, and $\tau$ is the time
pass between cycle $t$ and $t+1$. We can rewrite @eq-fitness as
$$
\frac{1}{\tau}\ln \frac{f_{t+1}^{(b)}}{f_{t}^{(b)}} = (s^{(b)} - \bar{s}_t).
$${#eq-logfreq}
@eq-logfreq separates the measurements---the barcode frequencies---from the
unobserved (sometimes referred to as latent) parameters we want to infer from
the data---the population mean fitness and the barcode relative fitness. This is
ultimately the functional form used in our inference pipeline. Therefore, the
relative fitness is computed by knowing the log frequency ratio of each barcode
throughout the growth-dilution cycles.

The presence of the neutral lineages facilitates the determination of the 
population mean fitness value $\bar{s}_t$. Since every relative fitness is 
determined relative to the neutral lineage that dominates the culture, we define
their fitness to be $s^{(n)} = 0$, where the superscript $(n)$ specifies their
neutrality. This means that @eq-logfreq for a neutral lineage takes the simpler
form
$$
\frac{1}{\tau}\ln \frac{f_{t+1}^{(n)}}{f_{t}^{(n)}} = - \bar{s}_t.
$${#eq-logfreq_neutral}
Therefore, we can use the data from these reference barcodes to directly infer
the value of the population mean fitness.

It is important to notice that the frequencies $f_{t}^{(b)}$ are not the allele
frequencies in the population (most of the culture is not sequenced since the
reference strain is not barcoded), but rather the relative frequencies in the
total number of sequencing reads. A way to conceptualize this subtle but
important point is to assume exponential growth in the *number of cells*
$N_t^{(b)}$ of the form
$$
N_{t+1}^{(b)} = N_{t}^{(b)} \mathrm{e}^{\lambda^{(b)}\tau},
$${#eq-ncells}
for every barcode $b$ with growth rate $\lambda^{(b)}$. However, when we
sequence barcodes, we do not directly measure the number of cells, but some
number of reads $r_t^{(b)}$ that map to barcode $b$. In the simplest possible
scenario, we assume
$$
r_{t}^{(b)} \propto N_{t}^{(b)},
$${#eq-r_to_n}
where, importantly, the proportionality constant depends on the total number of
reads for the library for cycle $t$, which might vary from library to library.
Therefore, to compare the number of reads between libraries at different time
points, we must normalize the number of reads to the same scale. The simplest
form is to define a relative abundance, i.e., a frequency with respect to the
total number of reads,
$$
f_{t}^{(b)} \equiv \frac{r_{t}^{(b)}}{\sum_{b'} r_{t}^{(b')}}.
$${#eq-r_to_n_norm}
This is the frequency @eq-fitness describes.

Our ultimate objective is to infer the relative fitness $s^{(m)}$ for each of
the $M$ relevant barcodes in the experiment. To do so, we account for the three
primary sources of uncertainty in our model:

1. Uncertainty in the determination of frequencies. Our model relates
frequencies between adjacent growth-dilution cycles to the fitness of the
corresponding strain. However, we do not directly measure frequencies. Instead,
our data for each barcode consists of a length $T$ vector of counts
$\underline{r}^{(b)}$ for each of the $T$ cycles in which the measurements were
taken.
2. Uncertainty in the value of the population mean fitness. We define neutral
lineages to have fitness $s^{(n)} = 0$, helping us anchor the value of the
population mean fitness $\bar{s}_t$ for each pair of adjacent growth cycles.
Moreover, we take this parameter as an empirical parameter to be obtained from
the data, meaning that we do not impose a functional form that relates
$\bar{s}_t$ to $\bar{s}_{t+1}$. Thus, we must infer the $T-1$ values of this
population mean fitness with their uncertainty that must be propagated to the
value of the mutants' relative fitness.
3. Uncertainty in each of the mutants' fitness values.

To account for all these sources of uncertainty in a principled way, in the next
section, we develop a Bayesian inference pipeline.