### 2023-09-26 doc/_bayesian_inference.qmd
small fixes to section

---
editor:
    render-on-save: true
bibliography: references.bib
csl: ieee.csl
---
## Bayesian inference {#sec-bayesian_inference}

As defined in @sec-fitness_model, our ultimate objective is to infer the vector
of relative fitness values
$$
\underline{s}^M = (s^{(1)}, s^{(2)}, \ldots, s^{(M)})^\dagger,
$${#eq-fitness_mut_vec}
where $^\dagger$ indicates the transpose. Our data consists of an $T \times B$
matrix $\underline{\underline{R}}$, where $B = M + N$ is the number of unique
barcodes given by the sum of the number of unique, relevant barcodes we care
about, $M$, and the number of unique neutral barcodes, $N$, and $T$ is the
number of growth cycles where measurements were taken. The data matrix is then
of the form
$$
\underline{\underline{R}} = \begin{bmatrix}
- & \underline{r}_1 & - \\
- & \underline{r}_2 & - \\
 & \vdots & \\
- & \underline{r}_T & - \\
\end{bmatrix},
$${#eq-R_mat}
where each row $\underline{r}_t$ is a $B$-dimensional array containing the raw
barcode counts at cycle $t$. We can further split each vector $\underline{r}_t$
into two vectors of the form
$$
\underline{r}_t = \begin{bmatrix}
\underline{r}_t^{N} \\
\underline{r}_t^{M}
\end{bmatrix},
$${#eq-r-vec_split}
i.e., the vector containing the neutral lineage barcode counts
$\underline{r}_t^{N}$ and the corresponding vector containing the mutant barcode
counts $\underline{r}_t^{M}$. Following the same logic, matrix
$\underline{\underline{R}}$ can be split into two matrices as
$$
\underline{\underline{R}} = \left[ 
\underline{\underline{R}}^N \; \underline{\underline{R}}^M
\right],
$${#eq-R-mat_split}
where $\underline{\underline{R}}^N$ is a $T \times N$ matrix with the barcode
reads time series for each neutral lineage and $\underline{\underline{R}}^M$ is
the equivalent $T \times M$ matrix for the non-neutral lineages.

Our objective is to compute the joint probability distribution for all relative
fitness values given our data. We can express this joint posterior distribution
using Bayes theorem as
$$
\pi(\underline{s}^M \mid \underline{\underline{R}}) = \frac{
\pi(\underline{\underline{R}} \mid \underline{s}^M) 
\pi(\underline{s}^M)}
{\pi(\underline{\underline{R}})},
$${#eq-bayes_obj}
where hereafter $\pi(\cdot)$ defines a probability density function, unless
otherwise stated. When defining our statistical model, we need not to focus on
the denominator on the right-hand side of @eq-bayes_obj. Thus, we can write
$$
\pi(\underline{s}^M \mid \underline{\underline{R}}) \propto
\pi(\underline{\underline{R}} \mid \underline{s}^M) 
\pi(\underline{s}^M).
$${#eq-bayes_obj_propto}
However, when implementing the model computationally, the normalization constant
on the right-hand side of @eq-bayes_obj must be computed. This can be done from
the definition of the model via an integral of the form
$$
\pi(\underline{\underline{R}}) = \int d^M \underline{s}^M
\pi(\underline{\underline{R}} \mid \underline{s}^M) 
\pi(\underline{s}^M),
$${#eq-bayes_evidence}
also known as a marginalization integral. Hereafter, differentials of the form
$d^n$ imply a $n$-dimensional integral.

Although @eq-bayes_obj and @eq-bayes_obj_propto seem simple enough, recall that
@eq-fitness relates barcode frequency values and the population mean fitness to
the mutant relative fitness. Therefore, we must include these nuisance
parameters as part of our inference problem. To include these nuisance
parameters, let 
$$
\underline{\bar{s}}_T = (\bar{s}_1, \bar{s}_2, \ldots, \bar{s}_{T-1})^\dagger,
$${#eq-pop_fitness_vec}
be the vector containing the $T-1$ population mean fitness we compute from the
$T$ time points where measurements were taken. We have $T-1$ since the value
of any $\bar{s}_t$ requires cycle numbers $t$ and $t+1$. Furthermore, let the
matrix $\underline{\underline{F}}$ be a $T \times B$ matrix containing all
frequency values. As with @eq-R-mat_split, we can split
$\underline{\underline{F}}$ into two matrices of the form
$$
\underline{\underline{F}} = \left[ 
\underline{\underline{F}}^N \; \underline{\underline{F}}^M
\right],
$${#eq-F-mat_split}
to separate the corresponding neutral and non-neutral barcode frequencies. With
these nuisance variables in hand, the full inference problem we must solve takes
the form
$$
\pi(
    \underline{s}^M, \underline{\bar{s}}_T, \underline{\underline{F}} \mid
    \underline{\underline{R}}
) \propto
\pi(
    \underline{\underline{R}} \mid
    \underline{s}^M, \underline{\bar{s}}_T, \underline{\underline{F}}
)
\pi(
    \underline{s}^M, \underline{\bar{s}}_T, \underline{\underline{F}}
).
$${#eq-bayes_full}
To recover the marginal distribution over the non-neutral barcodes relative
fitness values, we can numerically integrate out all nuisance parameters, i.e.,
$$
\pi(\underline{s}^M \mid \underline{\underline{R}}) =
\int d^{T-1}\underline{\bar{s}}_T
\int d^{B}\underline{f}_1 \cdots
\int d^{B}\underline{f}_T
\;
\pi(
    \underline{s}^M, \underline{\bar{s}}_T, \underline{\underline{F}} \mid
    \underline{\underline{R}}
).
$${#eq-bayes_full_marginal}

### Factorizing the posterior distribution {#seq-split_posterior}

The left-hand side of @eq-bayes_full is extremely difficult to work with.
However, we can take advantage of the structure of our inference problem to
rewrite it in a more manageable form. Specifically, the statistical dependencies
of our observations and latent variables allow us to factorize the joint
distribution into the product of multiple conditional distributions. To gain
some intuition about this factorization, let us focus on the inference of the
population mean fitness values $\underline{\bar{s}}_T$. @eq-logfreq_neutral
relates the value of the population mean fitness to the neutral lineage
frequencies and nothing else. This suggests that when writing the posterior for
these population mean fitness parameters, we should be able to condition it only
on the neutral lineage frequency values, i.e., $\pi(\underline{\bar{s}}_T \mid
\underline{\underline{F}}^N)$. We point the reader to @sec-bayes_def for the
full mathematical details on this factorization. For our purpose here, it
suffices to say we can rewrite the joint probability distribution as a product
of conditional distributions of the form
$$
\pi(
    \underline{s}^M, \underline{\bar{s}}_T, \underline{\underline{F}} \mid
    \underline{\underline{R}}
) =
\pi(
    \underline{s}^M \mid \underline{\bar{s}}_T, \underline{\underline{F}}^M
)
\pi(
    \underline{\bar{s}}_T \mid \underline{\underline{F}}^N
)
\pi(\underline{\underline{F}} \mid \underline{\underline{R}}).
$${#eq-split_posterior}
Written in this form, @eq-split_posterior captures the three sources of
uncertainty listed in @sec-fitness_model in each term. Starting from right to
left, the first term on the right-hand side of @eq-split_posterior accounts for
the uncertainty when inferring the frequency values given the barcode reads. The
second term accounts for the uncertainty in the values of the mean population
fitness at different time points. The last term accounts for the uncertainty in
the parameter we care about---the mutants' relative fitnesses. We refer the
reader to @sec-bayes_def for an extended description of the model with specific
functional forms for each term on the left-hand side of @eq-split_posterior as
well as the extension of the model to account for multiple experimental
replicates or hierarchical genotypes.

### Variational Inference

One of the technical challenges to the adoption of Bayesian methods is the
analytical intractability of integrals such as that of @eq-bayes_full_marginal.
Furthermore, even though efficient Markov Chain Monte Carlo (MCMC) algorithms
such as Hamiltonian Montecarlo can numerically perform this integration
[@betancourt2017], the dimensionality of the problem in @eq-split_posterior
makes an MCMC-based approach prohibitively slow.

To overcome this computational limitation, in this work, we rely on the recent
development of the automatic differentiation variational inference algorithm
(ADVI) [@kucukelbir2016]. Briefly, when performing ADVI, our target posterior
distribution $\pi(\theta \mid \underline{\underline{R}})$, where $\theta =
(\underline{s}^M, \underline{\bar{s}}_T, \underline{\underline{F}})$, is
replaced by an approximate posterior distribution $q_\phi(\theta)$, where $\phi$
fully parametrizes the approximate distribution. As further explained in
@sec-vi_primer, the numerical integration problem is replaced by an optimization
problem of the form
$$
q^*_\phi(\theta) = \min _\phi
D_{KL}(
    q_\phi(\theta) \lvert \lvert
    \pi(\theta \mid \underline{\underline{R}})
),
$${#eq-vi_objective}
where $D_{KL}$ is the Kulback-Leibler divergence. In other words, the
complicated high-dimensional numerical integration problem is transformed into a
much simpler problem of finding the value of the parameters $\phi$ such that
@eq-vi_objective is satisfied as best as possible within some finite computation
time. Although to compute @eq-vi_objective, we require the posterior
distribution we are trying to approximate $\pi(\theta \mid
\underline{\underline{R}})$, it can be shown that maximizing the so-called
evidence lower bound (ELBO) [@kingma2014]---equivalent to minimizing the
variational free energy [@buckley2017]---is mathematically equivalent to
performing the optimization prescribed by @eq-vi_objective. We direct the reader
to @sec-vi_primer for a short primer on variational inference.

This work is accompanied by the Julia library `BayesFitness.jl` that makes use
of the implementation of both MCMC-based integration as well as ADVI
optimization to numerically approximate the solution of @eq-bayes_full_marginal
within the Julia ecosystem [@ge2018].