### 2023-09-26 doc/_bayes_priors.qmd
finish prior definition section

---
editor:
    render-on-save: true
bibliography: references.bib
csl: ieee.csl
---

### Defining prior probabilities

One aspect commonly associated---in both positive and negative ways---to
Bayesian analysis is the definition of prior probabilities. On the one hand, the
naive textbook version of Bayesian analysis defines the prior as encoding the
information we have about the inference in question before acquiring any data.
This is the "ideal" use of priors that, whenever possible, should be
implemented. On the other hand, for most practitioners of Bayesian statistics in
the age of big data, the definition of prior becomes a tool to ensure the
convergence of sampling algorithms such as MCMC [@gelman2017]. However, for our
particular problem, although we deal with large amounts of data (inferences can
be made for > 10K barcodes over multiple time points, resulting in > 100K
parameters), each barcode has very little data, as they are measured only once
per time point over < 10 growth-dilution cycles. Furthermore, it is incredibly
challenging to understand the noise sources related to culturing conditions, DNA
extraction, library preparation, etc., and encode them into reasonable
prior distributions.

Empirically, our approach for this work defined the priors based solely on the
neutral lineage data, as they represent the only repeated measurements of a
single genotype in our experimental design. We acknowledge that defining the
priors after observing the data might be considered an incoherent inference.
However, as expressed by @gelman2017

> Incoherence is an unavoidable aspect of much real-world data analysis; and,
indeed, one might argue that as scientists we learn the most from the anomalies
and reassessments associated with episodes of incoherence.

With this in mind, we leave it to the reader to judge the selection of priors.
Furthermore, the software package associated with this work, `BayesFitness.jl`,
is written so that users can experiment with different prior selection criteria
that fit their needs. We strongly advocate that statistics should not be done in
a black-box fit-all tool mindset but rather as a formal way to encode the
assumptions behind the analysis, subject to constructive criticism. With this
philosophical baggage behind us, let us now focus on how the priors used for
this work were selected.

#### Naive neutral lineage-based priors

For the base model presented in this work, the user-defined prior parameters
include the following:

- Prior on population mean fitness (one per pair of adjacent time points)
$$
\bar{s}_t \sim \mathcal{N}(\mu_{\bar{s}_t}, \sigma_{{\bar{s}_t}}).
$${#eq-prior_meanfit_mean}

- Prior on standard deviation associated with neutral lineages likelihood
function (one per pair of adjacent time points)
$$
\sigma_t \sim \log\mathcal{N}(\mu_{\sigma_t}, \sigma_{\sigma_t}).
$${#eq-prior_meanfit_var}

- Prior on relative fitness (one per non-neutral barcode)
$$
s^{(m)} \sim \mathcal{N}(\mu_{s^{(m)}}, \sigma_{s^{(m)}}).
$${#eq-prior_fit_mean}

- Prior on standard deviation associated with non-neutral lineages likelihood
function (one per non-neutral barcode)
$$
\sigma^{(m)} \sim \log\mathcal{N}(\mu_{\sigma^{(m)}}, \sigma_{\sigma^{(m)}})
$${#eq-prior_fit_var}

The `BayesFitness.jl` package includes a function `naive_prior` within the
`stats` module. This function utilizes the data from the neutral lineages to
determine some of the prior parameters to facilitate the inference algorithm's
numerical convergence. In particular, it defines the population mean fitness
parameter $\mu_{\bar{s}_t}$ as
$$
\mu_{\bar{s}_t} = \frac{1}{N} \sum_{n=1}^N 
-\ln\left( \frac{r_{t+1}^{(n)}}{r_t^{(n)}} \right),
$${#eq-prior_meanfit_mean_data}
where $N$ is the number of neutral lineages and $r_t^{(n)}$ is the number of
neutral lineages. In other words, it defines the mean of the prior distribution
as the mean of what one naively would compute from the neutral lineages,
discarding cases where the ratio diverges because the denominator $r_t^{(n)} =
0$. For the variance parameter, we chose a value $\sigma_{\bar{s}_t} = 0.05$.

Furthermore, the `naive_prior` function defines the mean of the variance
parameter as the standard deviation of the log frequency ratios for the neutral
lineages, i.e.,
$$
\mu_{\sigma_t} = \sqrt{
    \operatorname{Var}\left(\frac{r_{t+1}^{(n)}}{r_t^{(n)}}\right)
},
$${#eq-prior_meanfit_var_data}
where $\operatorname{Var}$ is the sample variance. This same value was utilized
for the mean of the non-neutral barcode variance $\mu_{\sigma^{(m)}}$. While we
assigne the corresponding variances to be $\sigma_{\sigma_t} =
\sigma_{\sigma^{(m)}} = 1$.