### 2023-10-04 doc/_multienv_results.qmd

---
editor:
    render-on-save: true
bibliography: references.bib
csl: ieee.csl
---

## Fitness inference on multiple environments{#sec-multienv}

One of the benefits of Bayesian methods is the intrinsic error propagation
embedded in the mathematical framework. For example, the uncertainty on the
value of the population mean fitness values $\underline{\bar{s}}_T$ is
propagated to the relative fitness of a particular strain $s^{(b)}$ since we
defined a joint posterior distribution over all parameters as fully expressed in
@eq-bayes_full. This natural error propagation can help us with the experimental
design schematized in @fig-03(A). Here, rather than performing growth-dilution
cycles in the same environment, the cells are diluted into a different 
environment. Thus, the uncertainty on the fitness estimate for the previous
environment must be propagated to that of the next one.

To validate the extension of our statistical model to this scenario, @fig-03(B)
shows the trajectory of the log frequency ratios between adjacent time points.
The different colored regions correspond to the different environments. For this
simulation, the growth rate of Environment 2 was set to be, on average, half of
the average growth rate in Environment 1. Equivalently, the growth rate in
Environment 3 was set to be, on average, twice the average growth rate in
Environment 1. @fig-03(C-E) show the correspondence between the simulation
ground truth and the inferred fitness values, where the error bars represent the
68\% credible region. @fig-03(F) summarizes the performance of our inference
pipeline by showing the empirical cumulative distribution functions for the
absolute value of the ground truth fitness value z-score within the posterior
distribution. This plot shows that, overall, $\approx 75\%$ of inferred mean
values fall within one standard deviation of the ground truth. For completeness,
@fig-03(G) shows the posterior predictive checks for a few example barcodes.

![**Multi-environment fitness inference.** (A) Schematic of the simulated
experimental design where growth-dilution cycles are performed into different
environments for each cycle. (B) log frequency rations between adjacent time
points. Darker lines represent the neutral barcodes. The colors in the
background demark the corresponding environment, matching colors in (A).
Environment 2 is set to have, on average, half the growth rate of environment 1.
Likewise, environment 3 is set to have, on average, twice the growth rate of
environment 1. (C-E) Comparison between the ground truth fitness value from the
logistic-growth simulation and the inferred fitness value for each environment.
Gray error bars represent the 68\% posterior credible region. (F) The empirical
cumulative distribution function (ECDF) for the absolute z-score value of the
ground truth parameter value within the inferred fitness posterior distribution
for all fitness values (black line) and each environment individually (color
lines). (G) Examples of the posterior predictive checks for all neutral lineages
(upper left panel) and a subset of representative mutant lineages. Shaded
regions surrounding the data represent the 95\%, 68\%, and 5\% credible regions
for the data. The reported errors above the plot represent the 68\% credible
region on the mutant relative fitness marginal distribution. Background colors
match those of (A).](./figs/fig03){#fig-03}