### 2023-10-05 doc/_multienv_results.qmd
update section based on Madhav's comments

---
editor:
    render-on-save: true
bibliography: references.bib
csl: ieee.csl
---

## Fitness inference on multiple environments{#sec-multienv}

The fitness model in @eq-fitness relates nuisance parameters such as the
population mean fitness and the barcode frequencies to the relative fitness
parameter we want to infer from the data. These dependencies imply that
uncertainty on the estimates of these nuisance parameters influences the
inference of the relevant parameters. For example, imagine a scenario where the
neutral lineages data were incredibly noisy, leading to poor estimates of the
population mean fitness values $\underline{\bar{s}}_T$. Since the relative
fitness of any non-neutral barcode $s^{(m)}$ is determined with respect to these
neutral barcodes, not accounting for the lack of precision in the value of the
population mean fitness would result in misleading estimates of the accuracy
with which we determine the value of the parameter we care about. Thus,
propagating these sources of uncertainty in nuisance parameters is vital to
generate an unbiased estimate of the relevant information we want to extract
from the data. One of the benefits of Bayesian methods is the intrinsic error
propagation embedded in the mathematical framework. For our previous example,
the uncertainty on the value of the population mean fitness values is propagated
to the relative fitness of a non-neutral barcode since we defined a joint
posterior distribution over all parameters as fully expressed in @eq-bayes_full.

This natural error propagation can help us with the experimental design
schematized in @fig-03(A). Here, rather than performing growth-dilution cycles
in the same environment, the cells are diluted into a different environment.
Thus, the uncertainty on the fitness estimate for the previous environment must
be propagated to that of the next one. To validate the extension of our
statistical model to this scenario, @fig-03(B) shows the trajectory of the log
frequency ratios between adjacent time points. The different colored regions
correspond to the different environments. For this simulation, the growth rate
of Environment 2 was set to be, on average, half of the average growth rate in
Environment 1. Equivalently, the growth rate in Environment 3 was set to be, on
average, twice the average growth rate in Environment 1. @fig-03(C-E) show the
correspondence between the simulation ground truth and the inferred fitness
values, where the error bars represent the 68\% credible region. @fig-03(F)
summarizes the performance of our inference pipeline by showing the empirical
cumulative distribution functions for the absolute value of the ground truth
fitness value z-score within the posterior distribution. This plot shows that,
overall, $\approx 75\%$ of inferred mean values fall within one standard
deviation of the ground truth. For completeness, @fig-03(G) shows the posterior
predictive checks for a few example barcodes.

![**Multi-environment fitness inference.** (A) Schematic of the simulated
experimental design where growth-dilution cycles are performed into different
environments for each cycle. (B) log frequency rations between adjacent time
points. Darker lines represent the neutral barcodes. The colors in the
background demark the corresponding environment, matching colors in (A).
Environment 2 is set to have, on average, half the growth rate of environment 1.
Likewise, environment 3 is set to have, on average, twice the growth rate of
environment 1. (C-E) Comparison between the ground truth fitness value from the
logistic-growth simulation and the inferred fitness value for each environment.
Gray error bars represent the 68\% posterior credible region. (F) The empirical
cumulative distribution function (ECDF) for the absolute z-score value of the
ground truth parameter value within the inferred fitness posterior distribution
for all fitness values (black line) and each environment individually (color
lines). (G) Examples of the posterior predictive checks for all neutral lineages
(upper left panel) and a subset of representative mutant lineages. Shaded
regions surrounding the data represent the 95\%, 68\%, and 5\% credible regions
for the data. The reported errors above the plot represent the 68\% credible
region on the mutant relative fitness marginal distribution. Background colors
match those of (A).](./figs/fig03){#fig-03}