### 2023-10-11 doc/_abstract.qmd
add missing sections to paper

The tracking of lineage frequencies via DNA barcode sequencing enables the
quantification of microbial fitness. However, experimental noise coming from
biotic and abiotic sources complicates the computation of a reliable inference.
We present a Bayesian pipeline to infer relative microbial fitness from
high-throughput lineage tracking assays. Our model accounts for multiple sources
of noise and propagates uncertainties throughout all parameters in a systematic
way. Furthermore, using modern variational inference methods based on automatic
differentiation, we are able to scale the inference to a large number of unique
barcodes. We extend this core model to analyze multi-environment assays,
replicate experiments, and barcodes linked to genotypes. On simulations, our
method recovers known parameters within posterior credible intervals. This work
provides a generalizable Bayesian framework to analyze lineage tracking
experiments. The accompanying open-source software library enables the adoption
of principled statistical methods in experimental evolution.