### 2023-10-11 doc/_acknowledgements.qmd

We would like to thank Griffin Chure and Michael Betancourt for their helpful
advice and discussion. We would like to thank Karna Gowda, Spencer Farrell, and
Shaili Mathur for critical observations on the manuscript. MRM was supported by
the Schmidt Science Fellowship. MM was supported by The National Science
Foundation-Simons Center for Quantitative Biology at Northwestern University and
the Simons Foundation grant 597491. MM is a Simons Investigator.