### 2023-10-11 doc/_acknowledgements.qmd
update acknowledgements and affiliations

We would like to thank Griffin Chure and Michael Betancourt for their helpful
advice and discussion. We would like to thank Karna Gowda, Spencer Farrell, and
Shaili Mathur for critical observations on the manuscript. This work was
supported by the NIH/NIGMS, Genomics of rapid adaptation in the lab and in the
wild R35GM11816506 (MIRA grant), and the CZ Biohub investigator program. MRM was
supported by the Schmidt Science Fellowship. MM was supported by The National
Science Foundation-Simons Center for Quantitative Biology at Northwestern
University and the Simons Foundation grant 597491. MM is a Simons Investigator. 
DP is a CZ Biohub investigator.