### 2023-10-11 doc/_acknowledgements.qmd
finish adding all funding sources

We would like to thank Griffin Chure and Michael Betancourt for their helpful
advice and discussion. We would like to thank Karna Gowda, Spencer Farrell, and
Shaili Mathur for critical observations on the manuscript. This work was
supported by the NIH/NIGMS, Genomics of rapid adaptation in the lab and in the
wild R35GM11816506 (MIRA grant), the NIH, Unravelling mechanisms of tumor
suppression in lung cancer R01CA23434903, the NIH (PQ4), Quantitative and
multiplexed analysis of gene function in cancer in vivo R01CA23125303, the NIH,
Genetic Determinants of Tumor Growth and Drug Sensitivity in EGFR Mutant Lung
Cancer R01CA263715, the NIH, Dissecting the interplay between aging, genotype,
and the microenvironment in lung cancer U01AG077922, the NIH, Genetic dissection
of oncogenic Kras signaling R01CA230025, and the CZ Biohub investigator program.
MRM was supported by the Schmidt Science Fellowship. MM was supported by The
National Science Foundation-Simons Center for Quantitative Biology at
Northwestern University and the Simons Foundation grant 597491. MM is a Simons
Investigator. DP is a CZ Biohub investigator.