### 2023-11-17 doc/paper.qmd
add lua filters to split main text and SI text references in the PDF automatically

::: {.maintext}
# Introduction

{{< include _intro.qmd >}}

# Results

<!-- Experimental results -->
{{< include _experiment.qmd >}}

<!-- Fitness model -->
{{< include _fitness_model.qmd >}}

<!-- Bayesian inference -->
{{< include _bayesian_inference.qmd >}}

<!-- Single dataset -->
{{< include _single_results.qmd >}}

<!-- Multi-environment results -->
{{< include _multienv_results.qmd >}}

<!-- Hierarchical model for experimental replicates -->
{{< include _hierarchical_replicates_results.qmd >}}

<!-- Hierarchical model for genotypes-->
{{< include _hierarchical_genotypes_results.qmd >}}

# Discussion

{{< include _discussion.qmd >}}

# Acknowledgements

{{< include _acknowledgements.qmd >}}
:::

::: {.sitext}
# Supplementary Materials

<!-- Primer on Variational Bayes -->
{{< include _variational_inference_primer.qmd >}}

<!-- Bayesian inference pipeline -->
{{< include _bayes_freq.qmd >}}

{{< include _bayes_meanfit.qmd >}}

{{< include _bayes_mutfit.qmd >}}

<!-- Hierarchical models -->
{{< include _bayes_hierarchical.qmd >}}

<!-- Prior selection -->
{{< include _bayes_priors.qmd >}}

<!-- Posterior predictive checks -->
{{< include _ppc.qmd >}}

<!-- Logistic growth simulations -->
{{< include _logistic_sim.qmd >}}
:::