### 2023-12-04 doc/_fitseq2_sim.qmd
first draft of section

---
editor:
    render-on-save: true
bibliography: references.bib
csl: ieee.csl
---

## Comparison with `FitSeq2.0` {#sec-fitseq2}

For a complete comparison, we apply our inference pipeline to a simulated
dataset closely following the simulation pipeline from @li2023. This simulation
explicitly accounts for multiple noise sources captured in the added Gaussian
noise in @sec-logistic. Here, we list all the steps involved in this explicit
simulation, highlighting that the overall results are qualitatively very similar
to that of our simple logistic growth plus added Gaussian noise when computing
the relevant quantity of the log frequency ratio.

1. We initialized a population consisting of 1000 unique barcodes---50 neutral
barcodes and 950 mutant barcodes---where the initial number of cells with
barcode $i$ in the culture was sampled from a Gamma distribution as
$$
n_o^{(i)} \sim \text{Gamma}(20, 0.2)
$${#eq-fitseq_gamma_pop}

2. The Malthusian fitness (related but not exactly the relative fitness to be
inferred from the data) for non-neutral barcodes was then sampled from a skewed
normal distribution of the form
$$
\lambda^{(j)} \sim \text{Skew}\mathcal{N}(0, 0.225, 3)
$${#eq-fitseq_skewnormal}

3. To simulate the noisy growth, the number of cells underwent the following
steps:

**for** $t=2:T$ **do**

&nbsp;&nbsp;&nbsp;&nbsp; 
$n_{t,c}^{(i)} = \text{Poisson}(n_{t-1,c}^{(i)} e^{\lambda^{(i)}})$

**end for**,

where $T$ is the number of generations, and $c$ indexes the growth-dilution
cycle. We use $T=4$ for our simulation.

4. After the $T$ generations, cells were stochastically diluted back for the
next growth-dilution cycle as
$$
n_{0,c+1}^{(i)} \sim \text{Poisson}
\left(\frac{n_{T,c}^{(i)}}{2^{T}}\right).
$${#eq-fitseq_dilution}
This process was repeated $C=4$ times for a total of four growth-dilution
cycles.

5. To simulate the sampling noise when taking cells that will then be
subsequently processed to obtain barcode reads, we sampled
$n_{\text{sample},c}^{(i)}$ cells with barcode $i$ as
$$
n_{\text{sample},c}^{(i)} \sim 
\text{Poisson}(L \times N \times f_{c}^{(i)}),
$${#eq-fitseq_sample}
where $c$ indexes the growth cycle, $N$ is the total number of unique barcodes,
$f_{j}^{(i)}$ is the fraction of cells in the culture with barcode $i$ at the
end of cycle $j$, i.e.,
$$
f_{c}^{(i)} = \frac{n_{c}^{(i)}}
{\sum_{j=1}^N n_{c}^{(j)}},
$${#eq-fitseq_frac_cell}
and $L$ is the expected average number of cells per lineage. We used $L = 500$.

6. To simulate the noise due to PCR amplification, we iterate $c_{\text{PCR}} =
25$ times, stochastically doubling the number of barcodes with a Poisson sample
of the form

**for** $a=1:A$ **do**

&nbsp;&nbsp;&nbsp;&nbsp; 
$n_{\text{PCR},(a+1),c}^{(i)} \sim \text{Poisson}(2 n_{\text{PCR},(a),c}^{(i)})$

**end for**,

where $(a)$ indexes the amplification cycle, $A$ is the total number of cycles,
and $c$ indexes the growth-dilution cycle. We use $A=25$ for our simulation.For
$a = 1$, we use the number of cells sampled in @eq-fitseq_sample.

7. To simulate the noise due to the sampling step involved during the sequencing
of the amplicon library, the number of reads mapping to barcode $i$ at the end
of cycle $c$ used in the inference was sampled as 
$$
n_{\text{seq},c}^{(i)} \sim 
\text{Poisson}(R \times N \times f_{\text{PCR},c}^{(i)}),
$${#eq-fitseq_seq}
where $R$ is the expected number of reads per barcode (R=100 for our
simulation), $N$ is the number of total barcodes, and $f_{\text{PCR},c}^{(i)}$
is the fraction of barcodes of type $i$ in the final amplification cycle from
step 4.

@fig-SIX_fitseq2(A) and (B) show the simulation's frequency and log frequency
ratio trajectories, respectively. Qualitatively, these results are very similar
to those shown in Fig. 1 of @li2023. As done throughout the main text,
@fig-SIX_fitseq2(C) shows examples of the posterior predictive checks for the
neutral lineages and a few of the mutant lineages. We can see that our inference
pipeline can fit this simulated data for the examples shown. Furthermore,
@fig-SIX_fitseq2(D) compares the ground truth fitness values and the inferred
parameters for all mutant lineages. @fig-SIX_fitseq2 shows the comparison
between the inference with the software provided in @li2023 and our inference
pipeline. We note that although the values are highly correlated, the scale of
the axis is very different.

![**FitSeq2.0 simulation comparison**.(A) Frequency trajectories that represent
the raw data going into the inference. (B) Log frequency ratio between two
adjacent time points used by the inference pipeline. Darker lines represent the
neutral barcodes. These transformed data are much more noisy than the seemingly
smooth frequency trajectories. (C) Examples of the posterior predictive checks
for all neutral lineages (upper left panel) and a subset of representative
mutant lineages. Shaded regions represent the 95\%, 68\%, and 5\% credible
regions for the data. The reported errors above the plot represent the 68\%
credible region on the mutant relative fitness marginal distribution. (D)
Comparison between the ground truth fitness value from the logistic-growth
simulation and the inferred fitness value. Gray error bars represent the 68\%
posterior credible region for the relative fitness values. (E) Comparison
between fitness inferences as obtained using `FitSeq2.0` from @li2023 and our
Bayesian method. Vertical gray error bars represent the 68\% posterior credible
region for the relative fitness
values.](./figs/figSIX_fitseq2){#fig-SIX_fitseq2}