### 2023-12-11 doc/_kinsler.qmd
first draft of Kinsler et al SI section

---
editor:
    render-on-save: true
bibliography: references.bib
csl: ieee.csl
---

## Reanalyzing Kinsler et al., 2020 {#sec-kinsler}

The main text of this paper focuses only on simulation-based assessments of the
performance of our inference algorithm. This supplementary section shows how the
model performs on experimental data. In particular, we focus on a subset of the
data generated for [@kinsler2020]. In this paper, the fitness of a set of
mutants originally evolved in minimal glucose media was determined in multiple
environments using an experimental design described here. @fig-SIX_kinsler_data
shows the raw barcode frequency trajectories for four of such environments
measured in two biological replicates.

![**Kinsler et al., 2020 sample data**. Barcode frequency trajectories for
minimal glucose evolved mutants in four different environments. Each environment
was measured in two biological replicates (R1 and R2 on each panel). For
experimental details, see
[@kinsler2020].](./figs/figSIX_kinsler_data){#fig-SIX_kinsler_data}

### Analysis of individual replicates

As a first analysis, we perform inference assuming all replicates are
independent. @fig-SIX_kinsler_ppc_single shows the posterior predictive checks
for the neutral lineages (upper left plot on each panel) and some sample mutants
for both replicates. We can see that the inference works exceptionally well with
the data, capturing the variability. Although the main text inferences were made
on logistic growth simulations, as described in @sec-logistic, the results
in @sec-fitseq2 and this section show that our model does not depend on the
assumptions underlying such simulations. The only assumption that must be valid
is that the changes in the relative frequencies of each lineage follow an
exponential function.

![**Posterior predictive checks for Kinsler et al., 2020 independent
replicates**. Examples of the posterior predictive checks for all neutral
lineages (upper left panels) and a subset of representative mutant lineages for
all four environments shown in @fig-SIX_kinsler_data. Shaded regions from light
to dark represent the 95\%, 68\%, and 5\% credible regions. Listed fitness
values for mutant lineages represent the 68\% credible
region.](./figs/figSIX_kinsler_ppc_single){#fig-SIX_kinsler_ppc_single}

Furthermore, we can compare the replicate-to-replicate correlation for these
four environments with two biological replicates. The result of this comparison
is shown in @fig-SIX_kinsler_rep_single.

![**Replicate-to-replicate correlation**. Comparison of inferred fitness values
between experimental replicates when fitting each replicate individually for all
environments shown in @fig-SIX_kinsler_data. Gray error bars show the 68\%
credible region for this
inference.](./figs/figSIX_kinsler_rep_single){#fig-SIX_kinsler_rep_single}

### Hierarchical analysis of multiple replicates

As discussed in @sec-replicates, most experimental designs---including that of
[@kinsler2020]---include two or more biological replicate measurements.
Therefore, a natural way to analyze the data is by jointly inferring the fitness
value using a hierarchical model. @fig-SIX_kinsler_ppc_hier shows the posterior
predictive checks when fitting a hierarchical model to each of the four
environments shown in @fig-SIX_kinsler_data. Comparing this figure with
@fig-SIX_kinsler_ppc_single shows that the hierarchical model does not lose the
ability to capture the variability in the data compared to fitting each
replicate individually. However, as shown in @fig-SIX_kinsler_rep_hier, when
comparing the replicate-to-replicate correlation, as expected, the hierarchical
model does a better job.

![**Posterior predictive checks for Kinsler et al., 2020 hierarchical model**.
Examples of the posterior predictive checks for all neutral lineages (upper left
panels) and a subset of representative mutant lineages for all four environments
shown in @fig-SIX_kinsler_data. Shaded regions from light to dark represent the
95\%, 68\%, and 5\% credible regions. Listed fitness values for mutant lineages
represent the 68\% credible
region.](./figs/figSIX_kinsler_ppc_hier){#fig-SIX_kinsler_ppc_hier}

![**Replicate-to-replicate correlation**. Comparison of inferred fitness values
between experimental replicates when fitting a hierarchical model on the
biological replicates for all environments shown in @fig-SIX_kinsler_data. Gray
error bars show the 68\% credible region for this
inference.](./figs/figSIX_kinsler_rep_hier){#fig-SIX_kinsler_rep_hier}