### 2023-12-13 doc/_kinsler.qmd
extended analysis significantly

---
editor:
    render-on-save: true
bibliography: references.bib
csl: ieee.csl
---

## Reanalyzing Kinsler et al., 2020 {#sec-kinsler}

The main text of this paper focuses only on simulation-based assessments of the
performance of our inference algorithm. This supplementary section shows how the
model performs on experimental data. In particular, we focus on a subset of the
data generated for [@kinsler2020]. In this paper, the fitness of a set of
mutants originally evolved in minimal glucose media was determined in multiple
environments using an experimental design described in @sec-experiment.
@fig-SIX_kinsler_data shows the raw barcode frequency trajectories for four of
such environments measured in two biological replicates.

![**Kinsler et al., 2020 sample data**. Barcode frequency trajectories in four
different environments for mutants evolved in minimal glucose media. Each
environment was measured in two biological replicates (R1 and R2 on each panel).
For experimental details, see
[@kinsler2020].](./figs/figSIX_kinsler_data){#fig-SIX_kinsler_data}

### Analysis of individual replicates

As a first analysis, we perform inference assuming all replicates are
independent. @fig-SIX_kinsler_ppc_single shows the posterior predictive checks
for the neutral lineages (upper left plot on each panel) and some sample mutants
for both replicates. We can see that the inference works exceptionally well with
the data, capturing the variability. Although the main text inferences were made
on logistic growth simulations, as described in @sec-logistic, the results
in @sec-fitseq2 and this section show that our model does not depend on the
assumptions underlying such simulations. The only assumption that must be valid
is that the changes in the relative frequencies of each lineage follow an
exponential function.

![**Posterior predictive checks for Kinsler et al., 2020 independent
replicates**. Examples of the posterior predictive checks for all neutral
lineages (upper left panels) and a subset of representative mutant lineages for
all four environments shown in @fig-SIX_kinsler_data. Shaded regions from light
to dark represent the 95\%, 68\%, and 5\% credible regions. Listed fitness
values for mutant lineages represent the 68\% credible
region.](./figs/figSIX_kinsler_ppc_single){#fig-SIX_kinsler_ppc_single}

Furthermore, we can compare the replicate-to-replicate correlation for these
four environments with two biological replicates. The result of this comparison
is shown in @fig-SIX_kinsler_rep_single. We can see that there is a good degree
of correlation between replicates.

![**Replicate-to-replicate correlation for single-dataset model**. Comparison of
inferred relative fitness values for mutant strains between experimental
replicates when fitting each replicate individually for all environments shown
in @fig-SIX_kinsler_data. Gray error bars show the 68\% credible region for this
inference.](./figs/figSIX_kinsler_rep_single){#fig-SIX_kinsler_rep_single}

Given that we have two biological replicates of the same experiment, we can ask
how the difference between each replicate's inferred fitness values compares
with the individual estimates' uncertainty. In other words, we want to compare
how different the $x$- and $y$-axis values in @fig-SIX_kinsler_rep_single are
with the size of the shown error bars. @fig-SIX_kinsler_meandiff_single compares
the distribution of these magnitudes for all four environments. In each plot,
the blue line shows the empirical cumulative distribution function (ECDF) of the
distance between the mean fitness estimate for each replicate---computed as
$\sqrt{\left( \left\langle s_1^{(m)}\right\rangle - \left\langle
s_2^{(m)}\right\rangle \right)^2}$, where $\langle s_i^{(m)}\rangle$ is the mean
fitness estimate of mutant $m$ in replicate $i$---meaning the difference between
$x$- and $y$-axis in @fig-SIX_kinsler_rep_single. The orange and green lines
represent the ECDF for each replicate's posterior distribution standard
deviation, i.e., the ECDF of the size of the error bars in
@fig-SIX_kinsler_rep_single. From these plots, we can see that the distribution
of standard deviations is shifted to the left compared to the difference between
replicates. This implies that the uncertainty in the estimate is generally
larger than the difference between replicates one would obtain with a naive
point estimate.

![**Comparing the replicate-to-replicate deviations with within-batch noise
estimates for single dataset analysis**. Empirical cumulative distribution
functions of the distance between the mean fitness estimate for each replicate
(blue lines) and each replicate's posterior distribution standard deviation
(orange and green
line).](./figs/figSIX_kinsler_meandiff_single){#fig-SIX_kinsler_meandiff_single}

### Hierarchical analysis of multiple replicates

As discussed in @sec-replicates, most experimental designs---including that of
@kinsler2020---include two or more biological replicate measurements. Therefore,
a natural way to analyze the data is by jointly inferring the fitness value
using a hierarchical model. @fig-SIX_kinsler_ppc_hier shows the posterior
predictive checks when fitting a hierarchical model to each of the four
environments shown in @fig-SIX_kinsler_data. Comparing this figure with
@fig-SIX_kinsler_ppc_single shows that the hierarchical model does not lose the
ability to capture the variability in the data compared to fitting each
replicate individually. However, as shown in @fig-SIX_kinsler_rep_hier, the
hierarchical model does a better job when comparing the replicate-to-replicate
correlation, as expected. To emphasize this last point,
@fig-SIX_kinsler_rep_comb shows both inferences on the same plot---only the mean
of the posterior distribution is shown for visual clarity.

![**Posterior predictive checks for Kinsler et al., 2020 hierarchical model**.
Examples of the posterior predictive checks for all neutral lineages (upper left
panels) and a subset of representative mutant lineages for all four environments
shown in @fig-SIX_kinsler_data. Shaded regions from light to dark represent the
95\%, 68\%, and 5\% credible regions. Listed fitness values for mutant lineages
represent the 68\% credible
region.](./figs/figSIX_kinsler_ppc_hier){#fig-SIX_kinsler_ppc_hier}

![**Replicate-to-replicate correlation for hierarchical model**. Comparison of
inferred fitness values between experimental replicates when fitting a
hierarchical model on the biological replicates for all environments shown in
@fig-SIX_kinsler_data. Gray error bars show the 68\% credible region for this
inference.](./figs/figSIX_kinsler_rep_hier){#fig-SIX_kinsler_rep_hier}

![**Replicate-to-replicate correlation for both methods**. Comparison of
**mean** inferred fitness values between experimental replicates when fitting a
single-dataset model (blue points) vs. a hierarchical model on the biological
replicates (orange points) for all environments shown in @fig-SIX_kinsler_data.
Error bars are not included in the plot for visual
clarity](./figs/figSIX_kinsler_rep_comb){#fig-SIX_kinsler_rep_comb}

We can again compare the magnitude of the difference between replicates'
inferred mean fitness with the size of the corresponding error bars, as we did
with @fig-SIX_kinsler_meandiff_single. @fig-SIX_kinsler_meandiff_hier shows the
ECDF for the resulting difference between replicates mean fitness in blue as
well as the resulting standard deviation for the inferred hyper-fitness
parameter that connects both replicates in orange (see @sec-replicates for
description of this hyper-fitness parameter). Here, we see that the standard
deviation ECDF (orange line) is shifted more to the right than the mean
difference between replicates (blue line).

![**Comparing the replicate-to-replicate deviations with within-batch noise
estimates for hierarchical model analysis**. Empirical cumulative distribution
functions of the distance between the mean fitness estimate for each replicate
(blue lines) and each replicate's posterior distribution standard deviation
(orange and green
line).](./figs/figSIX_kinsler_meandiff_hier){#fig-SIX_kinsler_meandiff_hier}

Even though the difference between replicates is reduced when using a
hierarchical model, this does not come at the expense of higher uncertainty in
the estimate of the relative fitness. To show this is the case,
@fig-SIX_kinsler_std_comp compares the reported posterior distribution standard
deviations for the relative fitness when inferring each dataset individually
($x$-axis) with the standard deviation of the hyper-fitness parameter
($y$-axis). Most of the points fall below the identity line, demonstrating that
the uncertainty in the hyper-fitness parameter obtained by utilizing information
from both replicates is smaller than the equivalent uncertainty obtained with
one replicate at a time. In conclusion, the hierarchical model not only can
equally describe the data, as shown in @fig-SIX_kinsler_ppc_hier, but the
difference between replicates is significantly decreased, as shown in
@fig-SIX_kinsler_rep_comb, at the same time, the uncertainty in the parameters'
estimate is decreased, as shown in @fig-SIX_kinsler_std_comp.

![**Comparing estimates uncertainties between single-dataset and hierarchical
model**. Comparison of the reported standard deviations for the mutants'
relative fitness when inferring the relative fitness using one dataset at a time
($x$-axis) and the corresponding standard deviation for the mutants'
hyper-fitness parameter when utilizing both replicates simultaneously via a
hierarchical model ($y$-axis). Dashed line shows identity
line.](./figs/figSIX_kinsler_std_comp){#fig-SIX_kinsler_std_comp}