### 2023-12-20 doc/_discussion.qmd
Add explanation of fitness model dynamics and future directions

---
editor:
    render-on-save: true
bibliography: references.bib
csl: ieee.csl
---

Experimental evolution of microbial systems has dramatically advanced our
understanding of the basic principles of biological evolution [@kussell2013].
From questions related to the optimal fine-tuning of gene expression programs
[@Dekel2005], to the dimensionality, geometry, and accessibility of the adaptive
fitness landscape explored by these rapidly adapting populations
[@kinsler2020;@maeda2020], to the emergence of eco-evolutionary dynamics in a
long-term evolution experiment [@good2017]; for all of these and other cases,
the microbial experimental platform combined with high-throughput sequencing has
been essential to tackling these questions with empirical data. This exciting
research area promises to improve as new culturing technologies [@jagdish2022]
as well as more complex lineage barcoding schemes [@nguyenba2019a;@yang2022],
are adopted.

For this data-heavy field, properly accounting for the uncertainty in parameters
inferred from experiments is vital to ensure the conclusions drawn are reliable.
Bayesian statistics presents a principled way to quantify this uncertainty
systematically [@gelman2010]. Moreover, Bayesian analysis offers a more
natural way to interpret the role that probability theory plays when performing
data analysis compared to the often-misinterpreted frequentist methods
[@vanderplas2014]. Nevertheless, the technical challenges associated with
Bayesian analysis has limited its application. This is set to change as
recognition of the misuse of frequentist concepts such as the p-value is
receiving more attention [@nuzzo2014]. Moreover, advances in numerical methods
such as Hamiltonian Monte Carlo [@betancourt2017] and variational inference
[@kucukelbir2016] allows for complex Bayesian models to be fit to empirical
data.

In this paper, we present a computational pipeline to analyze lineage-tracking
time-series data for massive-parallel competition assays. More specifically, we
fit a Bayesian model to infer the fitness of multiple genotypes relative to a
reference [@kinsler2020;@ascensao2023]. The proposed model accounts for multiple
sources of uncertainty with proper error propagation intrinsic to Bayesian
methods. To scale the inference pipeline to large datasets with $> 10,000$
barcodes, we use the ADVI algorithm [@kucukelbir2016] to fit a variational
posterior distribution. The main difference between our method and previous
inference pipelines, such as @li2023, is that the present analysis provides
interpretable errors on the inferred fitness values. The reported uncertainty
intervals---known as credible regions---can be formally interpreted as capturing
the corresponding probability mass of finding the true value of the parameter
given the model, the prior information, and the data. Furthermore, minor
modifications to the structure of the statistical model presented in this work
allow for the analysis of different experimental designs, such as
growth-dilution cycles in different environments, joint analysis of multiple
experimental replicates of the same experiment via hierarchical models, and a
hierarchical model for multiple barcodes mapping to equivalent genotypes. We
validate our analysis pipeline on simulated datasets with known ground truth,
showing that the model fits the data adequately, capturing the ground truth
parameters within the posterior distribution.

It is important to highlight some of the consequences of the general
experimental design and the implicit assumptions within the proposed statistical
model to analyze the resulting data. First, the composition of the population is
such that the initial fraction of the population occupied by the barcoded
genotypes is small---usually >90\% of the initial population is the non-labeled
reference strain. This constraint is important as the fitness model used to fit
the time series data assumes that the tracked frequencies are $\ll 1$. Second,
when computing log frequency ratios, we can run into the issue of dividing by
zero. This is a common problem when dealing with molecular count data
[@lovell2020]. Our model gets around this issue by assuming that the frequency
of any barcode cannot be, but still can get arbitrarily close to, zero.
Therefore, we implicitly assume that no lineage goes extinct during the
experiment. Moreover, the statistical model directly accounts for the
uncertainty associated with having zero barcode counts, increasing the
corresponding uncertainty. Third, the models presented in this paper require the
existence of a labeled sub-population of barcoded reference strains. These
barcodes help determine the fitness baseline, as every fitness is quantified
with respect to this reference genotype. This experimental design constraint
facilitates the inference of the population mean fitness since most of the
culture---the unlabeled reference genotype---is not tracked. 
<span style="color: #912E27;">
Fourth, the fitness model in @eq-fitness does not assume any functional form for
the cellular growth dynamics. This description of the changes in the relative
frequency of the genotypes can approximate the dynamics for multiple growth
forms. For example, in @sec-fitseq2, in the supplementary materials, we apply
our inference pipeline to simulations like those presented in [@li2023] with
non-logistic growth models. Moreover, @sec-kinsler, in the supplementary
materials, reanalyzes experimental data from [@kinsler2020], where previous work
showed that the fitness advantage from some of the mutants results from a
shorter lag phase at the beginning of the growth cycle [@li2018]. Thus, our
method's inferred relative fitness is a coarse-grained quantity that integrates
the differences in fitness across the entire growth cycle as long as frequencies
and frequency changes are small enough to be approximated by @eq-fitness.
</span>
Finally, the presented statistical model assumes that relative fitness is solely
a constant of the environment and the genotype. Future directions of this work
could extend the fitness model to properly analyze data with time-varying or
frequency-dependent fitness values.

In total, the statistical model presented in this work and the software package
accompanying the paper allow for a principled way of quantifying the accuracy
with which we can extract relevant parametric information from large-scale
multiplexed fitness competition assays. Furthermore, the implementation of
Bayesian models and their fitting via automatic differentiation approaches opens
the gate to extend this type of formal analysis to the data-rich literature in
experimental evolution and other high-throughput technologies applications.