### 2023-12-31 doc/_time_scaling.qmd
Add time scaling section to paper

---
editor:
    render-on-save: true
bibliography: references.bib
csl: ieee.csl
---

## Computation time scaling {#sec-scaling}

The biggest advantage of the variational inference approach presented in this
work is the computation speedup compared to MCMC-based Bayesian approaches. This
allows for applying our principled Bayesian pipeline to datasets with many
unique barcodes or multiple datasets analyzed hierarchically.
@fig-SIX_bc_vs_time shows a linear scaling in the amount of time taken for the
ADVI algorithm to take 7,500 steps for a varying number of unique barcodes. This
figure shows that the time to fit larger models increases linearly with the
number of barcodes.

![**Scaling of computation time as a function of number of unique barcodes**.
Fitting a single-dataset model for 500 to 100,000 unique barcodes by taking
7,500 steps.](./figs/figSIX_bc_vs_time){#fig-SIX_bc_vs_time}