### 2024-02-26 doc/_author_summary.qmd
upload missing document modifications

---
editor:
    render-on-save: true
bibliography: references.bib
csl: ieee.csl
---

In this study, we present a novel Bayesian pipeline for analyzing DNA barcode
tracking sequencing data, addressing the challenge of accurately quantifying
competitive microbial fitness in the presence of experimental noise. Our method
uniquely contributes to understanding microbial evolutionary dynamics by
enabling reliable inference of the relative fitness of diverse microbial strains
from high-throughput lineage tracking assays. Our approach is distinct in its
ability to systematically account for and propagate uncertainties from various
noise sources throughout all inferred parameters. Furthermore, the
error-propagation quality of our Bayesian method allows us to extend the
inference pipeline to common dataset structures, such as jointly analyzing
multiple experimental replicates or accounting for multiple unique barcodes
mapping the equivalent genotypes. This comprehensive treatment of uncertainties
is crucial in experimental settings where noise can significantly influence the
results. Furthermore, we have optimized our pipeline for scalability, allowing
it to handle large numbers of unique barcodes effectively. This scalability is
essential for analyzing complex datasets typical in microbial fitness studies.