### 2024-02-26 doc/paper.qmd

::: {.maintext}

# Abstract {.unnumbered}

The tracking of lineage frequencies via DNA barcode sequencing enables the
quantification of microbial fitness. However, experimental noise coming from
biotic and abiotic sources complicates the computation of a reliable inference.
We present a Bayesian pipeline to infer relative microbial fitness from
high-throughput lineage tracking assays. Our model accounts for multiple sources
of noise and propagates uncertainties throughout all parameters in a systematic
way. Furthermore, using modern variational inference methods based on automatic
differentiation, we are able to scale the inference to a large number of unique
barcodes. We extend this core model to analyze multi-environment assays,
replicate experiments, and barcodes linked to genotypes. On simulations, our
method recovers known parameters within posterior credible intervals. This work
provides a generalizable Bayesian framework to analyze lineage tracking
experiments. The accompanying open-source software library enables the adoption
of principled statistical methods in experimental evolution.

# Author Summary {.unnumbered}

{{< include _author_summary.qmd >}}

# Introduction

{{< include _intro.qmd >}}

# Results

<!-- Experimental results -->
{{< include _experiment.qmd >}}

<!-- Fitness model -->
{{< include _fitness_model.qmd >}}

<!-- Bayesian inference -->
{{< include _bayesian_inference.qmd >}}

<!-- Single dataset -->
{{< include _single_results.qmd >}}

<!-- Multi-environment results -->
{{< include _multienv_results.qmd >}}

<!-- Hierarchical model for experimental replicates -->
{{< include _hierarchical_replicates_results.qmd >}}

<!-- Hierarchical model for genotypes-->
{{< include _hierarchical_genotypes_results.qmd >}}

# Discussion

{{< include _discussion.qmd >}}

# Materials and Methods {.unnumbered}

All synthetic data generation and custom scripts used in this work were stored
using Git version control. Code for analysis and figure generation can be found
on the GitHub repository (https://github.com/mrazomej/bayesian_fitness). The
accompanying software package `BarBay.jl` can be directly installed from the
`Julia` package repository, or via cloning the corresponding GitHub repository
(https://github.com/mrazomej/BarBay.jl)

# Acknowledgements {.unnumbered}

{{< include _acknowledgements.qmd >}}
:::

::: {.sitext}
# Supplementary Materials

<!-- Primer on Variational Bayes -->
{{< include _variational_inference_primer.qmd >}}

<!-- Bayesian inference pipeline -->
{{< include _bayes_freq.qmd >}}

{{< include _bayes_meanfit.qmd >}}

{{< include _bayes_mutfit.qmd >}}

<!-- Hierarchical models -->
{{< include _bayes_hierarchical.qmd >}}

<!-- Prior selection -->
{{< include _bayes_priors.qmd >}}

<!-- Posterior predictive checks -->
{{< include _ppc.qmd >}}

<!-- Logistic growth simulations -->
{{< include _logistic_sim.qmd >}}

<!-- Stochastic growth simulations -->
{{< include _fitseq2_sim.qmd >}}

<!-- Kinsler et al. 2020 -->
{{< include _kinsler.qmd >}}

<!-- Scaling as a function of number of barcodes -->
{{< include _time_scaling.qmd >}}
:::